---
title: "0_EDM_Vis_Raw"
author: "Kurt Ingeman"
date: "2/11/2020"
output:
  pdf_document: default
  html_document: default
---

## Examine raw and normalized series time data for samon recruitment and drivers


```{r include=FALSE, echo=FALSE, warning=FALSE}

rm(list = ls(all.names = TRUE)) 

library(rEDM)
library(tidyr)
library(ggplot2)
library(PNWColors)
library(gridExtra)
library(dplyr)
library(reshape2)

```


```{R include=FALSE, warning=FALSE}

load("Data/block_data.Rdata")

dat <- do.call("rbind", block_data) # omg I can't believe how long it took me to find the right  

```


```{r include=FALSE, warning=FALSE}

dat <-  dat %>% filter(year > 1956)

annual <- dat %>% 
  group_by(year) %>% 
  mutate(rec.ann = sum(rec, na.rm = T)) %>% 
  mutate(rec3.ann = sum(rec3, na.rm = T)) %>% 
  mutate(rec4.ann = sum(rec4, na.rm = T)) %>% 
  mutate(rec5.ann = sum(rec5, na.rm = T))

norm_ann <- dat %>% 
  group_by(year) %>% 
  mutate(rec.ann.n = mean(rec_n, na.rm = T)) %>% 
  mutate(rec3.ann.n = mean(rec3_n, na.rm = T)) %>% 
  mutate(rec4.ann.n = mean(rec4_n, na.rm = T)) %>% 
  mutate(rec5.ann.n = mean(rec5_n, na.rm = T))

```

### Total recruitment (summed across all MPGs)
```{r echo=FALSE, warning=FALSE}

# start with raw salmon numbers in a given brood year
r <- ggplot(annual, aes(x = year, y = rec.ann)) +
  geom_line(color = "black", size = .2) +
  geom_point(shape = 21, size = 2.5, fill = "white", colour = "black") +
  theme_bw() +
  labs(x = "Year", y = "Total recruitment from brood year")

print(r)

# next with normalized salmon numbers in a given brood year
r1 <- ggplot(norm_ann, aes(x = year, y = rec.ann.n)) +
  geom_line(color = "black", size = .2) +
  geom_point(shape = 21, size = 2.5, fill = "white", colour = "black") +
  theme_bw() +
  labs(x = "Year", y = " Mean of normalized recruitment")

# the high values in later years (when total rec is not high) tells one that the normalized values of some stocks is skewing late

```
<style>
div.blue { background-color:#e6f0ff; border-radius: 5px; padding: 20px;}
</style>
<div class = "blue">
### Takeaways
#### - Total recruitment (summed across all MPGs) follows well known env regimes 
#### - Exhibits rough downward trend 
</div>

### Salmon Recruitment by MPG
```{r echo=FALSE, warning=FALSE}
df <- dat %>% 
  select(mpg,stk,year,rec,rec3,rec4,rec5) %>% 
  gather(cohort, number, rec:rec5)

df_n <- dat %>% 
  select(mpg,stk,year,rec_n, rec3_n, rec4_n, rec5_n) %>% 
  gather(cohort, number, rec_n:rec5_n)

df_ann <- df %>% 
  group_by(year, cohort) %>% 
  mutate(ann.total = sum(number, na.rm = T))

df_ann_n <- df_n %>% 
  group_by(year, cohort) %>% 
  mutate(ann.mean = mean(number, na.rm = T))

df_mpg <- df %>% 
  group_by(year, mpg) %>% 
  summarise(ann.total = sum(number, na.rm = T))

df_c_m <- df %>% 
  group_by(year, cohort, mpg) %>% 
  summarise(ann.total = sum(number, na.rm = T))

# df_cohort <- df %>% 
#  group_by(year, Cohort, mpg) %>% 
#  summarise(ann.total = sum(Number, na.rm = T))

pal1 <- pnw_palette(name="Sunset2",type="discrete")
pal2 <- pnw_palette(name="Shuksan2",type="discrete")
pal3 <- pnw_palette(name="Starfish",type="discrete")
pal4 <- pnw_palette(name="Bay",type="discrete")

r2 <- ggplot(df_mpg, aes(x = year, y = ann.total)) +
  geom_line(color = "black", size = .2) +
#  geom_point(shape = 21, size = 1, fill = "white", colour = "black") +
  theme_bw() +
  labs(x = "Year", y = "Total recruitment from brood year") +
  facet_wrap(~mpg, scales = "free", nrow = 2)
print (r2)
```
<style>
div.blue { background-color:#e6f0ff; border-radius: 5px; padding: 20px;}
</style>
<div class = "blue">
#### - MPGs share common high/low years (eg 1998-1999) but of different magnitudes
#### - But longer term trends vary among MPGS (eg Imnaha vs All Salmon 1960-1970)
</div>

### Salmon recruitment colored by MPG
```{r echo=FALSE, warning=FALSE}
r3 <- ggplot(df_mpg, aes(x = year, y = ann.total, color = mpg)) +
  geom_line(size = .5) +
  theme_bw() +
  labs(x = "Year", y = "Total recruitment from brood year") +
  scale_color_manual(values = pal4)
print (r3)
```

<style>
div.blue { background-color:#e6f0ff; border-radius: 5px; padding: 20px;}
</style>
<div class = "blue">
#### - All MPGs declining overall
#### - Imnaha and Upper Salmon historical strongholds
#### - Increased synchrony/decreased portfolio of variability in given year (!)
</div>

### Total Recruitment and individual age cohorts
```{r echo=FALSE, warning=FALSE}

r4 <- ggplot(df_ann) +
  geom_line(aes(x = year, y = ann.total, color = cohort),  size = .5) +
  theme_bw() +
  scale_color_manual(values = pal4) +
  labs(x = "Year", y = "Total + cohort returns from same broood year")
print (r4)

```
<style>
div.blue { background-color:#e6f0ff; border-radius: 5px; padding: 20px;}
</style>
<div class = "blue">
#### - 5 year old returners a declining proportion even, esp in the good year of the mid-aughts
</div>


### Total Recruitment and individual age cohorts by MPG
```{r echo=FALSE, warning=FALSE}
r4a <- ggplot(df_c_m) +
  geom_line(aes(x = year, y = ann.total, color = cohort),  size = .5) +
  theme_bw() +
  scale_color_manual(values = pal4) +
  labs(x = "Year", y = "Total + cohort returns from same broood year") +
  facet_wrap(~mpg, scales = "free")
print (r4a)
```
<style>
div.blue { background-color:#e6f0ff; border-radius: 5px; padding: 20px;}
</style>
<div class = "blue">
#### - Pattern of fewe age 5 not driven by any one MPG but fairly robust pattern
</div>

### Normalized recruitment for every stock in each MPG
```{r echo=FALSE, warning=FALSE}
norm_rec <- df_ann_n %>% filter(cohort == "rec_n")

r5 <- ggplot(norm_rec, aes(x = year, y = number, group = interaction(mpg,stk), color = interaction(mpg,stk))) +
  geom_line() +
  facet_wrap(~mpg) +
  theme_bw() +
  theme(legend.position = "none") +
    labs(x = "Year", y = "Total returns by individual stock")
print (r5)

```
<style>
div.blue { background-color:#e6f0ff; border-radius: 5px; padding: 20px;}
</style>
<div class = "blue">
#### - Stocks within MPG's largely covary (with exception of North Fork(!) in Upper Salmon)
</div>


### Plot total returns against aligned predictors; facet by predictor
```{r echo=FALSE, warning=FALSE}

predictors <- dat %>% 
  select(mpg, stk, year, harvest, up48.apr.2, up48.nov.2, pdo.win.2, npgo.yr.2, hatch.spring.2, SRKW.JKLpods.4, pinn.CSLmales.4, pinn.SSLaerial.2, rec, rec3, rec4, rec5, rec_n) %>% 
  gather(cohort, number, rec:rec5)

predict_norm <- dat %>% 
  select(mpg, stk, year, harvest, up48.apr.2, up48.nov.2, pdo.win.2, npgo.yr.2, hatch.spring.2, SRKW.JKLpods.4, pinn.CSLmales.4, pinn.SSLaerial.2, rec_n, rec3_n, rec4_n, rec5_n) %>% 
  gather(cohort, number, rec_n:rec5_n)

raws <- predictors %>% 
  filter(cohort == "rec") %>% 
  group_by(year) %>% 
  summarise(mean.harvest = mean(harvest, na.rm = T), 
            ann.total = sum(number, na.rm = T))

r6 <- ggplot(raws) +
  geom_line(color = "black", size = .2, 
            aes(x = year, y = ann.total)) +
    geom_point(aes(x = year, y = ann.total), 
            shape = 21, size = 2.5, fill = "white", colour = "black")

r6a <- r6 +
    geom_line(aes(x = year, y = mean.harvest*50000), size = .5, color = "#a8554e") +
  theme_bw() +
  scale_y_continuous(sec.axis = sec_axis(~./50000, name = "Prop Harvested")) +
  labs(y = "Total Recruitment", x = "Year") +
  ggtitle("HARVEST") + 
   theme(plot.title = element_text(vjust = -1, hjust = .5)) 



norms <- predict_norm %>% 
  filter(cohort == "rec_n") %>% 
  group_by(year) %>% 
  summarise(ann.mean = mean(number, na.rm = T), 
            rec_norm = mean(number, na.rm = T),
            CSL = mean(pinn.CSLmales.4, na.rm = T), 
            SSL = mean(pinn.SSLaerial.2, na.rm = T), 
            up.apr = mean(up48.apr.2, na.rm = T), 
            up.nov = mean(up48.nov.2, na.rm = T), 
            pdo = mean(pdo.win.2, na.rm = T), 
            npgo = mean(npgo.yr.2, na.rm = T), 
            hatch = mean(hatch.spring.2))

r6n <- ggplot(norms) +
  geom_line(color = "black", size = .2, 
            aes(x = year, y = ann.mean)) +
    geom_point(aes(x = year, y = ann.mean), 
            shape = 21, size = 2.5, fill = "white", colour = "black")

r6b <- r6n +
    geom_line(aes(x = year, y = -pdo), size = .5, color = "#b0cbe7") +
#    geom_line(aes(x = year, y = up.nov), size = .3, color = "brown") +
  theme_bw() +
  scale_y_continuous(sec.axis = sec_axis(~., name = "PDO (neg)")) +
  labs(y = "Mean Norm Rec", x = "Year") +
  ggtitle("ENV (PDO)") + 
   theme(plot.title = element_text(vjust = -1, hjust = .5)) 

r6c <- r6n +
    geom_line(aes(x = year, y = hatch), size = .5, color = "#eba07e") +
#    geom_line(aes(x = year, y = up.nov), size = .3, color = "brown") +
  theme_bw() +
  scale_y_continuous(sec.axis = sec_axis(~., name = "Hatchery releases")) +
  labs(y = "Mean Norm Rec", x = "Year")  +
  ggtitle("HATCHERY") + 
   theme(plot.title = element_text(vjust = -1, hjust = .5)) 

 
r6d <- r6n +
    geom_line(aes(x = year, y = CSL), size = .5, color = "#5d74a5") +
#    geom_line(aes(x = year, y = up.nov), size = .3, color = "brown") +
  theme_bw() +
  scale_y_continuous(sec.axis = sec_axis(~., name = "CSL")) +
  labs(y = "Mean Norm Rec", x = "Year")  +
  ggtitle("CSL") + 
   theme(plot.title = element_text(vjust = -1, hjust = .5)) 


grid.arrange(r6d, r6b, r6c, r6a)

```
<style>
div.blue { background-color:#e6f0ff; border-radius: 5px; padding: 20px;}
</style>
<div class = "blue">
##### Lots to unpack here, first note that salmon numbers are mean of normalized stocks (except for in the harvest panel, where it is Total Recruitment 
#### - CSL at total population scale does not seem to be driving declines
#### - Closely correlated with PDO but total population bufffered against early events?
#### - Hatchery releases coincide with major declines consistent with replacement hypothesis
#### - Proportion harvested has tracked population incredibly well ... until recently(!)
</div>


### Zoom into CSL to see any stronger individual MPG correlations
```{r echo=FALSE, warning=FALSE}

norms_mpg <- predict_norm %>% 
  filter(cohort == "rec_n") %>% 
  group_by(year, mpg) %>% 
  summarise(ann.mean = mean(number, na.rm = T), 
            rec_norm = mean(number, na.rm = T),
            CSL = mean(pinn.CSLmales.4, na.rm = T), 
            SSL = mean(pinn.SSLaerial.2, na.rm = T), 
            up.apr = mean(up48.apr.2, na.rm = T), 
            up.nov = mean(up48.nov.2, na.rm = T), 
            pdo = mean(pdo.win.2, na.rm = T), 
            npgo = mean(npgo.yr.2, na.rm = T), 
            hatch = mean(hatch.spring.2, na.rm = T))

r7 <- ggplot(norms_mpg) +
  geom_line(color = "black", size = .2, 
            aes(x = year, y = ann.mean)) +
    geom_point(aes(x = year, y = ann.mean), 
            shape = 21, size = 2.5, fill = "white", colour = "black") +
  facet_wrap(~mpg, scales ="free")

r7a <- r7 +
    geom_line(aes(x = year, y = CSL), size = .3, color = "red") +
  theme_bw() +
  scale_y_continuous(sec.axis = sec_axis(~., name = "CSL")) +
  labs(y = "Total Recruitment", x = "Year")
print(r7a)

```
<style>
div.blue { background-color:#e6f0ff; border-radius: 5px; padding: 20px;}
</style>
<div class = "blue">
#### - Nope
</div>


### Zoom into harvest to look at individual MPG correlations
```{r echo=FALSE, warning=FALSE}

harv_mpg <- predictors %>% 
  filter(cohort == "rec") %>% 
  filter(!mpg == "Lower Snake") %>% 
  group_by(year, mpg) %>% 
  summarise(mean.harvest = mean(harvest, na.rm = T), ann.total = sum(number, na.rm = T))


r8 <- ggplot(harv_mpg) +
  geom_line(color = "black", size = .2, 
            aes(x = year, y = ann.total)) +
    geom_point(aes(x = year, y = ann.total), 
            shape = 21, size = 2.5, fill = "white", colour = "black") +
    geom_line(aes(x = year, y = mean.harvest*10000), size = .3, color = "red") +
  theme_bw() +
  scale_y_continuous(sec.axis = sec_axis(~./10000, name = "Prop Harvested")) +
  labs(y = "Total Recruitment", x = "Year") +
  facet_wrap(~mpg, nrow = 2, scales = "free")

print(r8)

```

<style>
div.blue { background-color:#e6f0ff; border-radius: 5px; padding: 20px;}
</style>
<div class = "blue">
#### - Seems that managers are unreasonably good at matching F to abundance, even at the MPG-level ... fishy
#### - Need to dig up meta data on harvest proportion 
</div>


## Now, look at covariance among predictors
```{r, echo=FALSE, warning=FALSE}
cov <- predictors %>% 
  filter(cohort == "rec") %>% 
  select(-c(cohort, up48.apr.2, npgo.yr.2, pinn.SSLaerial.2))

cov$year <- as.numeric(cov$year)

c1 <- pairs(predictors[,3:10], lower.panel = NULL, pch = 21, size = .2)
print(c1)

```
<style>
div.blue { background-color:#e6f0ff; border-radius: 5px; padding: 20px;}
</style>
<div class = "blue">
#### - A little hard to see, maybe use corrplot for colors
</div>

### Correlation heatmap of selected variables 
```{r echo=FALSE}

cormat <- round(cor(cov[,3:10], use = "na.or.complete"),2)
reorder_cormat <- function(cormat){
# Use correlation between variables as distance
dd <- as.dist((1-cormat)/2)
hc <- hclust(dd)
cormat <-cormat[hc$order, hc$order]
}

cormat <- reorder_cormat(cormat)

mel_cor <- melt(cormat)


# Get lower triangle of the correlation matrix
  get_lower_tri<-function(cormat){
    cormat[upper.tri(cormat)] <- NA
    return(cormat)
  }
  # Get upper triangle of the correlation matrix
  get_upper_tri <- function(cormat){
    cormat[lower.tri(cormat)]<- NA
    return(cormat)
  }
  
upper_tri <- get_upper_tri(cormat)

melted_cormat <- reshape2::melt(upper_tri, na.rm = TRUE)

c2 <- ggplot(data = melted_cormat, aes(Var2, Var1, fill = value)) +
 geom_tile(color = "white")+
 scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
   midpoint = 0, limit = c(-1,1), space = "Lab", 
   name="Pearson\nCorrelation") +
  theme_minimal() + 
  theme(axis.text.x = element_text(angle = 45, vjust = 1, 
    size = 12, hjust = 1), 
    axis.text.y = element_text(size= 12))+
 coord_fixed()

c2 + 
  geom_text(aes(Var2, Var1, label = value), color = "black", size = 4) +
  theme(axis.title.x = element_blank(),
  axis.title.y = element_blank(),
  panel.grid.major = element_blank(),
  panel.border = element_blank(),
  panel.background = element_blank(),
  axis.ticks = element_blank(),
  legend.justification = c(1, 0),
  legend.position = c(0.6, 0.7),
  legend.direction = "horizontal")+
  guides(fill = guide_colorbar(barwidth = 7, barheight = 1,
                title.position = "top", title.hjust = 0.5)) +
  scale_x_discrete(labels = c("PDO","Harvest", "Recruits", "Upwelling", "SRKW", "Hatcheries", "Year", "CSL")) +
  scale_y_discrete(labels = c("PDO","Harvest", "Recruits", "Upwelling", "SRKW", "Hatcheries", "Year", "CSL"))

```
<style>
div.blue { background-color:#e6f0ff; border-radius: 5px; padding: 20px;}
</style>
<div class = "blue">
#### - Strong positive correlations: Year X CSL, Hatcheries X CSL, Hatcheries X SRKW (!) 
#### - Strong negative correlation: PDO X CSL, PDO X Recruits
#### - Weak but notable: Recruits are positively linked to CSL but negative to SRKW
</div>

### Time to follow up simple correlations with some CCM results ...
<br><br><br><br>
