---
title: "0_EDM_Vis_Raw"
author: "Kurt Ingeman"
date: "2/11/2020"
output:
  pdf_document: default
  html_document: default
---

## Examine raw and normalized series time data for samon recruitment and drivers


```{r include=FALSE, echo=FALSE, warning=FALSE}

rm(list = ls(all.names = TRUE)) 

library(rEDM)
library(tidyr)
library(ggplot2)
library(PNWColors)
library(gridExtra)
library(dplyr)
library(reshape2)
library(patchwork)
library(pracma)
library(viridis)


```


```{R include=FALSE, warning=FALSE}

load("~/Github/Coordinated-Recoveries-GitHub/CORE EDM analysis/Data/Rdata/block_data.Rdata")

dat <- do.call("rbind", block_data) # omg I can't believe how long it took me to unlist this data object ... in 26 characters of code 
dat[dat$mpg == "Lower Snake", ]
# names(dat)
```


### Compile annual totals for the entire SRSS
```{r include=FALSE, warning=FALSE}

dat <-  dat %>% filter(year > 1956) 

## toy data for moving average
time <- 1:200
a <- runif(200, 100, 110)
b <- movavg(a, 10, type = "s")
c <- movavg(a, 25, type = "s")

x <- data.frame(time,a,b,c)

ggplot() +
  geom_point(color = "black", aes(x = time, y = a)) +
  geom_line(color = "red", aes(x = time, y = b)) +
  geom_line(color = "blue", aes(x = time, y = c))
## end toy

annual <- dat %>% 
  group_by(year) %>% 
  mutate(rec.ann = sum(rec, na.rm = T)) %>% 
  mutate(rec3.ann = sum(rec3, na.rm = T)) %>% 
  mutate(rec4.ann = sum(rec4, na.rm = T)) %>% 
  mutate(rec5.ann = sum(rec5, na.rm = T)) %>% 
  mutate(year2 = year + 2) %>%
  mutate(spawn_ann_nat = sum(na.omit(spawn_nat))) %>%
  mutate(movAveRec5 = movavg(rec.ann, 5, type = "s")) %>%
  mutate(spawn_ann_tot = sum(na.omit(spawn_tot))) %>% 
  ungroup()

norm_ann <- dat %>%   
  group_by(year) %>% 
  mutate(rec.ann.n = mean(rec_n, na.rm = T)) %>% 
  mutate(rec3.ann.n = mean(rec3_n, na.rm = T)) %>% 
  mutate(rec4.ann.n = mean(rec4_n, na.rm = T)) %>% 
  mutate(rec5.ann.n = mean(rec5_n, na.rm = T))

```

## Spawner-recruit relationship at multiple scales
### First, entire ESU

```{r}
require(scales)

# s
ggplot(data = annual) +
  geom_point(aes(x = spawn_ann_tot, y = rec.ann), shape = 21, size = 2.5, fill = "white", colour = "black") +
  scale_x_continuous(labels = scales::comma) +
  scale_y_continuous(labels = scales::comma) +
 geom_smooth(aes(x = spawn_ann_tot, y = rec.ann), method = "loess", formula = y ~ x, span = 10)

fig_spawn_rec <- ggplot(data = annual) +
  geom_point(aes(x = spawn_ann_tot, y = rec.ann, fill = year), shape = 21, size = 7, alpha = .5) +
  scale_x_continuous(labels = scales::comma) +
  scale_y_continuous(labels = scales::comma) +
  scale_fill_viridis(direction = -1) +
  labs(title = "Spawner-Recruit SRSS", x =  "Total Spawners (incl hatchery)", y = "Total Recruits") 

ggsave(filename =         
 "Figs/Manuscript/FigA_spawner_recruit.pdf", plot = fig_spawn_rec, width = 9, height = 5, units = "in") 
print(fig_spawn_rec)

```

### Next, spawner-recruit at MPG scale use facets
```{r}
# MPG pallette
pal=pnw_palette("Shuksan2",5, type = "discrete")

annual_mpg <- dat %>% 
  group_by(year, mpg) %>% 
  summarise(rec_mpg_ann = sum(rec, na.rm = T), 
            spawn_mpg_nat = sum(na.omit(spawn_nat)), 
            spawn_mpg_tot = sum(na.omit(spawn_tot)))

levels(annual_mpg$mpg)

# reorder levels for color
annual_mpg$mpg <- factor(annual_mpg$mpg, levels = c("Lower Snake", "Imnaha", "South Fork Salmon", "Middle Fork Salmon", "Upper Salmon"))

ggplot(data = annual_mpg) +
  geom_point(aes(x = spawn_mpg_tot, y = rec_mpg_ann, fill = mpg), shape = 21, size = 5, alpha = .7) +
  labs(title = "Spawner-Recruit SRSS", x =  "Total Spawners (incl hatchery)", y = "Total Recruits") +
  scale_x_continuous(labels = scales::comma) +
  scale_y_continuous(labels = scales::comma) +
  scale_fill_manual(values = rev(pal)) +
  theme(legend.position = "blank") +
  facet_wrap(~mpg)

fig_spawn_rec_mpg <- ggplot(data = annual_mpg) +
  geom_point(aes(x = spawn_mpg_tot, y = rec_mpg_ann, fill = mpg), shape = 21, size = 5, alpha = .7) +
  labs(title = "Spawner-Recruit SRSS", x =  "Total Spawners (incl hatchery)", y = "Total Recruits") +
  scale_x_continuous(labels = scales::comma) +
  scale_y_continuous(labels = scales::comma) +
  scale_fill_manual(values = rev(pal)) +
  theme(legend.position = "blank") +
  facet_wrap(~mpg, scales = "free")

ggsave(filename =         
 "Figs/Manuscript/FigB_spawner_recruit_MPG.pdf", plot = fig_spawn_rec_mpg , width = 9, height = 5, units = "in") 

print(fig_spawn_rec_mpg)


```




## Time series of whole SRSS recruitment (with PDO)

``` {r include=FALSE, warning=FALSE}

ggplot() +
  geom_point(data = yearly, color = "black", aes(x = year2, y = total_rec)) +
  geom_line(data = yearly, color = "red", aes(x = year2, y = movAveRec5))

ts_SRSS_pdo <- ggplot(yearly) +
    theme_minimal() +
  geom_rect(aes(xmin=1976, xmax=1998, ymin=0, ymax=Inf), 
            fill = "lightgray", alpha = 0.1) +
  geom_rect(aes(xmin=2002, xmax=2007, ymin=0, ymax=Inf), 
              fill = "lightgray", alpha = 0.9) +
  geom_rect(aes(xmin=2013, xmax=2020, ymin=0, ymax=Inf), 
           fill = "lightgray", alpha = 0.9) +  
  geom_line(aes(x = year2, y = total_rec), color = "black", size = .2) +
  geom_point(aes(x = year2, y = total_rec), shape = 21, size = 2.5, fill = "white", colour = "black") +
   geom_line(aes(x = year2, y = movAveRec5), color = "red", size = .2) +
  labs(title = "SRSS Returns across PDO regimes", x =  "First Ocean Year", y = "Total SRSS Returns") +
  xlim(1957, 2020)

ggsave(filename =         
 "Figs/Manuscript/FigA_salmon_time_series.pdf", plot = ts_SRSS_pdo, width = 5, height = 3, units = "in") 
 
print(ts_SRSS_pdo)

```


### Total recruitment (summed across all MPGs)
```{r echo=FALSE, warning=FALSE}

# start with raw salmon numbers in a given brood year
r <- ggplot(annual) +
  geom_line(aes(x = year2, y = rec.ann), color = "black", size = .2) +
  geom_point(aes(x = year2, y = rec.ann), shape = 21, size = 2.5, fill = "white", colour = "black") +
  theme_bw() +
  labs(x = "First Ocean Year", y = "Recruitment") +
  xlim(1957, 2020)

print(r)

# next with normalized salmon numbers in a given brood year
r1 <- ggplot(norm_ann, aes(x = year, y = rec.ann.n)) +
  geom_line(color = "black", size = .2) +
  geom_point(shape = 21, size = 2.5, fill = "white", colour = "black") +
  theme_bw() +
  labs(x = "Year", y = " Mean of normalized recruitment")

# the high values in later years (when total rec is not high) tells one that the normalized values of some stocks is skewing late

```

### Takeaways
#### - Total recruitment (summed across all MPGs) follows well known env regimes 
#### - Exhibits rough downward trend 

### Next, add time series of putative predictors (preds, harvest, hatcheries, environment)

```{r}
sst_win <- read.csv("Data/ersstArc.win.csv") %>% 
  filter(year > 1956)
sst_spr <- read.csv("Data/ersstArc.spr.csv") %>%
  filter(year > 1956)
sst <- full_join(sst_win, sst_spr, by = "year") %>% 
  mutate(year1 = year + 1) %>% 
  mutate(year2 = year + 2)


  
# Shuksan2 = rbind(c('#5d74a5', '#b0cbe7', '#fef7c7', '#eba07e', '#a8554e'))

rSST <- ggplot(sst) +
  geom_line(aes(x= year, y = ersstArc.spr), 
            color = "#5d74a5", size = .3) +
  geom_point(aes(x = year, y = ersstArc.spr), 
             shape = 21, size = 2.5, fill = "white", colour = "#5d74a5") +
   geom_line(aes(x = year, y = ersstArc.win), 
            color = "#a8554e", size = .3) +
  geom_point(aes(x = year, y = ersstArc.win), 
             shape = 22, size = 2.5, fill = "white", colour = "#a8554e") + 
  theme_bw() +
  labs(x = "Year", y = "SST") +
  xlim(1957, 2020)

r/rSST 

salm_ts <- dat %>% 
  group_by(year) %>% 
  summarise(rec.ann = sum(rec, na.rm = T)) %>% 
  mutate(year1 = year + 1) %>% 
  mutate(year2 = year + 2) %>% 
  mutate(period = case_when(
          year1 > 1998 ~ "late",
          year1 > 1977 & year <= 1997 ~ "mid",
          year1 <= 1976 ~ "early"))

ts <- left_join(salm_ts, sst, by = c("year2" = "year"))

ggplot(ts) +
  geom_point(aes(x = ersstArc.spr, y = rec.ann, color = period))


other_ts <- read.csv("Data/CORE2020_Pred_Harv_Hatch.csv")

all_ts <- left_join(salm_ts, other_ts, by = "year")

all_ts <- all_ts %>% 
  mutate_all(funs(as.numeric),2:34)


```

```{r}
## Plot min/max normalized data

normalize <- function(x) {
return ((x - min(x)) / (max(x) - min(x)))
}

norm_ts <- all_ts %>% 
    mutate(funs(normalize),2:34)

## Convert from wide to long for plotting
norm_long <- norm_ts %>% 
  gather(key = "vars", value = "counts", 2:34)

## split time series by variable 
ts_preds <- norm_long %>% 
  filter(grepl('pred.', vars))

ts_KW <- norm_long %>% 
  filter(grepl('KW.', vars))

ts_SSL <- norm_long %>% 
  filter(grepl('SSL.', vars))

ts_harv <- norm_long %>% 
  filter(grepl('harv.', vars))

ts_hatch <- norm_long %>% 
  filter(grepl('hatch.', vars))

as1 <- ggplot(norm_ts, aes(x = year, y = rec.ann)) +
  geom_line(color = "black", size = .2) +
  geom_point(shape = 21, size = 2.5, fill = "white", colour = "black") +
  theme_bw() +
  labs(x = "", y = "Total recruitment from brood year")

as2 <- ggplot(ts_KW, aes(x = year, y = counts, color = vars)) +
  geom_line() +
  theme_bw() +
  labs(x = "", y = "Killer Whale abundance")

as3 <- ggplot(ts_harv, aes(x = year, y = counts, color = vars)) +
  geom_line() +
  theme_bw() +
  labs(x = "", y = "Harvest")

as4 <- ggplot(ts_hatch, aes(x = year, y = counts, color = vars)) +
  geom_line() +
  theme_bw() +
  labs(x = "", y = "Hatcheries")

as1/as2/as3/as4
```


### Salmon Recruitment by MPG
```{r echo=FALSE, warning=FALSE}
df <- dat %>% 
  select(mpg,stk,year,rec,rec3,rec4,rec5) %>% 
  gather(cohort, number, rec:rec5)

df_n <- dat %>% 
  select(mpg,stk,year,rec_n, rec3_n, rec4_n, rec5_n) %>% 
  gather(cohort, number, rec_n:rec5_n)

df_ann <- df %>% 
  group_by(year, cohort) %>% 
  mutate(ann.total = sum(number, na.rm = T))

df_ann_n <- df_n %>% 
  group_by(year, cohort) %>% 
  mutate(ann.mean = mean(number, na.rm = T))

df_mpg <- df %>% 
  group_by(year, mpg) %>% 
  summarise(ann.total = sum(number, na.rm = T))

df_c_m <- df %>% 
  group_by(year, cohort, mpg) %>% 
  summarise(ann.total = sum(number, na.rm = T))

# df_cohort <- df %>% 
#  group_by(year, Cohort, mpg) %>% 
#  summarise(ann.total = sum(Number, na.rm = T))

pal1 <- pnw_palette(name="Sunset2",type="discrete")
pal2 <- pnw_palette(name="Shuksan2",type="discrete")
pal3 <- pnw_palette(name="Starfish",type="discrete")
pal4 <- pnw_palette(name="Bay",type="discrete")

r2 <- ggplot(df_mpg, aes(x = year, y = ann.total)) +
  geom_line(color = "black", size = .2) +
#  geom_point(shape = 21, size = 1, fill = "white", colour = "black") +
  theme_bw() +
  labs(x = "Year", y = "Total recruitment from brood year") +
  facet_wrap(~mpg, scales = "free", nrow = 2)
print (r2)
```

#### - MPGs share common high/low years (eg 1998-1999) but of different magnitudes
#### - But longer term trends vary among MPGS (eg Imnaha vs All Salmon 1960-1970)


### Salmon recruitment colored by MPG
```{r echo=FALSE, warning=FALSE}
r3 <- ggplot(df_mpg, aes(x = year, y = ann.total, color = mpg)) +
  geom_line(size = .5) +
  theme_bw() +
  labs(x = "Year", y = "Total recruitment from brood year") +
  scale_color_manual(values = pal4)
print (r3)
```

#### - All MPGs declining overall
#### - Imnaha and Upper Salmon historical strongholds
#### - Increased synchrony/decreased portfolio of variability in given year (!)


### Total Recruitment and individual age cohorts
```{r echo=FALSE, warning=FALSE}

r4 <- ggplot(df_ann) +
  geom_line(aes(x = year, y = ann.total, color = cohort),  size = .5) +
  theme_bw() +
  scale_color_manual(values = pal4) +
  labs(x = "Year", y = "Total + cohort returns from same broood year")
print (r4)

```

#### - 5 year old returners a declining proportion even, esp in the good year of the mid-aughts


### Total Recruitment and individual age cohorts by MPG
```{r echo=FALSE, warning=FALSE}
r4a <- ggplot(df_c_m) +
  geom_line(aes(x = year, y = ann.total, color = cohort),  size = .5) +
  theme_bw() +
  scale_color_manual(values = pal4) +
  labs(x = "Year", y = "Total + cohort returns from same broood year") +
  facet_wrap(~mpg, scales = "free")
print (r4a)
```

#### - Pattern of fewer age 5 not driven by any one MPG but fairly robust pattern


### Normalized recruitment for every stock in each MPG
```{r echo=FALSE, warning=FALSE}
norm_rec <- df_ann_n %>% filter(cohort == "rec_n")

r5 <- ggplot(norm_rec, aes(x = year, y = number, group = interaction(mpg,stk), color = interaction(mpg,stk))) +
  geom_line() +
  facet_wrap(~mpg) +
  theme_bw() +
  theme(legend.position = "none") +
    labs(x = "Year", y = "Total returns by individual stock")
print (r5)

```

#### - Stocks within MPG's largely covary (with exception of North Fork(!) in Upper Salmon)



### Plot total returns against aligned predictors; facet by predictor
```{r echo=FALSE, warning=FALSE}

predictors <- dat %>% 
  select(mpg, stk, year, harvest, up48.apr.2, up48.nov.2, pdo.win.2, npgo.yr.2, hatch.spring.2, SRKW.JKLpods.4, pinn.CSLmales.4, pinn.SSLaerial.2, rec, rec3, rec4, rec5, rec_n) %>% 
  gather(cohort, number, rec:rec5)

predict_norm <- dat %>% 
  select(mpg, stk, year, harvest, up48.apr.2, up48.nov.2, pdo.win.2, npgo.yr.2, hatch.spring.2, SRKW.JKLpods.4, pinn.CSLmales.4, pinn.SSLaerial.2, rec_n, rec3_n, rec4_n, rec5_n) %>% 
  gather(cohort, number, rec_n:rec5_n)

raws <- predictors %>% 
  filter(cohort == "rec") %>% 
  group_by(year) %>% 
  summarise(mean.harvest = mean(harvest, na.rm = T), 
            ann.total = sum(number, na.rm = T))

r6 <- ggplot(raws) +
  geom_line(color = "black", size = .2, 
            aes(x = year, y = ann.total)) +
    geom_point(aes(x = year, y = ann.total), 
            shape = 21, size = 2.5, fill = "white", colour = "black")

r6a <- r6 +
    geom_line(aes(x = year, y = mean.harvest*50000), size = .5, color = "#a8554e") +
  theme_bw() +
  scale_y_continuous(sec.axis = sec_axis(~./50000, name = "Prop Harvested")) +
  labs(y = "Total Recruitment", x = "Year") +
  ggtitle("HARVEST") + 
   theme(plot.title = element_text(vjust = -1, hjust = .5)) 



norms <- predict_norm %>% 
  filter(cohort == "rec_n") %>% 
  group_by(year) %>% 
  summarise(ann.mean = mean(number, na.rm = T), 
            rec_norm = mean(number, na.rm = T),
            CSL = mean(pinn.CSLmales.4, na.rm = T), 
            SSL = mean(pinn.SSLaerial.2, na.rm = T), 
            up.apr = mean(up48.apr.2, na.rm = T), 
            up.nov = mean(up48.nov.2, na.rm = T), 
            pdo = mean(pdo.win.2, na.rm = T), 
            npgo = mean(npgo.yr.2, na.rm = T), 
            hatch = mean(hatch.spring.2))

r6n <- ggplot(norms) +
  geom_line(color = "black", size = .2, 
            aes(x = year, y = ann.mean)) +
    geom_point(aes(x = year, y = ann.mean), 
            shape = 21, size = 2.5, fill = "white", colour = "black")

r6b <- r6n +
    geom_line(aes(x = year, y = -pdo), size = .5, color = "#b0cbe7") +
#    geom_line(aes(x = year, y = up.nov), size = .3, color = "brown") +
  theme_bw() +
  scale_y_continuous(sec.axis = sec_axis(~., name = "PDO (neg)")) +
  labs(y = "Mean Norm Rec", x = "Year") +
  ggtitle("ENV (PDO)") + 
   theme(plot.title = element_text(vjust = -1, hjust = .5)) 

r6c <- r6n +
    geom_line(aes(x = year, y = hatch), size = .5, color = "#eba07e") +
#    geom_line(aes(x = year, y = up.nov), size = .3, color = "brown") +
  theme_bw() +
  scale_y_continuous(sec.axis = sec_axis(~., name = "Hatchery releases")) +
  labs(y = "Mean Norm Rec", x = "Year")  +
  ggtitle("HATCHERY") + 
   theme(plot.title = element_text(vjust = -1, hjust = .5)) 

 
r6d <- r6n +
    geom_line(aes(x = year, y = CSL), size = .5, color = "#5d74a5") +
#    geom_line(aes(x = year, y = up.nov), size = .3, color = "brown") +
  theme_bw() +
  scale_y_continuous(sec.axis = sec_axis(~., name = "CSL")) +
  labs(y = "Mean Norm Rec", x = "Year")  +
  ggtitle("CSL") + 
   theme(plot.title = element_text(vjust = -1, hjust = .5)) 


grid.arrange(r6d, r6b, r6c, r6a)

```

##### Lots to unpack here, first note that salmon numbers are mean of normalized stocks (except for in the harvest panel, where it is Total Recruitment 
#### - CSL at total population scale does not seem to be driving declines
#### - Closely correlated with PDO but total population bufffered against early events?
#### - Hatchery releases coincide with major declines consistent with replacement hypothesis
#### - Proportion harvested has tracked population incredibly well ... until recently(!)


### Zoom into CSL to see any stronger individual MPG correlations
```{r echo=FALSE, warning=FALSE}

norms_mpg <- predict_norm %>% 
  filter(cohort == "rec_n") %>% 
  group_by(year, mpg) %>% 
  summarise(ann.mean = mean(number, na.rm = T), 
            rec_norm = mean(number, na.rm = T),
            CSL = mean(pinn.CSLmales.4, na.rm = T), 
            SSL = mean(pinn.SSLaerial.2, na.rm = T), 
            up.apr = mean(up48.apr.2, na.rm = T), 
            up.nov = mean(up48.nov.2, na.rm = T), 
            pdo = mean(pdo.win.2, na.rm = T), 
            npgo = mean(npgo.yr.2, na.rm = T), 
            hatch = mean(hatch.spring.2, na.rm = T))

r7 <- ggplot(norms_mpg) +
  geom_line(color = "black", size = .2, 
            aes(x = year, y = ann.mean)) +
    geom_point(aes(x = year, y = ann.mean), 
            shape = 21, size = 2.5, fill = "white", colour = "black") +
  facet_wrap(~mpg, scales ="free")

r7a <- r7 +
    geom_line(aes(x = year, y = CSL), size = .3, color = "red") +
  theme_bw() +
  scale_y_continuous(sec.axis = sec_axis(~., name = "CSL")) +
  labs(y = "Total Recruitment", x = "Year")
print(r7a)

```

#### - Nope

### Zoom into harvest to look at individual MPG correlations
```{r echo=FALSE, warning=FALSE}

harv_mpg <- predictors %>% 
  filter(cohort == "rec") %>% 
  filter(!mpg == "Lower Snake") %>% 
  group_by(year, mpg) %>% 
  summarise(mean.harvest = mean(harvest, na.rm = T), ann.total = sum(number, na.rm = T))


r8 <- ggplot(harv_mpg) +
  geom_line(color = "black", size = .2, 
            aes(x = year, y = ann.total)) +
    geom_point(aes(x = year, y = ann.total), 
            shape = 21, size = 2.5, fill = "white", colour = "black") +
    geom_line(aes(x = year, y = mean.harvest*10000), size = .3, color = "red") +
  theme_bw() +
  scale_y_continuous(sec.axis = sec_axis(~./10000, name = "Prop Harvested")) +
  labs(y = "Total Recruitment", x = "Year") +
  facet_wrap(~mpg, nrow = 2, scales = "free")

print(r8)

```


#### - Seems that managers are unreasonably good at matching F to abundance, even at the MPG-level ... fishy
#### - Need to dig up meta data on harvest proportion 



## Now, look at covariance among predictors
```{r, echo=FALSE, warning=FALSE}
cov <- predictors %>% 
  filter(cohort == "rec") %>% 
  select(-c(cohort, up48.apr.2, npgo.yr.2, pinn.SSLaerial.2))

cov$year <- as.numeric(cov$year)

c1 <- pairs(predictors[,3:10], lower.panel = NULL, pch = 21, size = .2)
print(c1)

```

#### - A little hard to see, maybe use corrplot for colors


### Correlation heatmap of selected variables 
```{r echo=FALSE}

cormat <- round(cor(cov[,3:10], use = "na.or.complete"),2)
reorder_cormat <- function(cormat){
# Use correlation between variables as distance
dd <- as.dist((1-cormat)/2)
hc <- hclust(dd)
cormat <-cormat[hc$order, hc$order]
}

cormat <- reorder_cormat(cormat)

mel_cor <- melt(cormat)


# Get lower triangle of the correlation matrix
  get_lower_tri<-function(cormat){
    cormat[upper.tri(cormat)] <- NA
    return(cormat)
  }
  # Get upper triangle of the correlation matrix
  get_upper_tri <- function(cormat){
    cormat[lower.tri(cormat)]<- NA
    return(cormat)
  }
  
upper_tri <- get_upper_tri(cormat)

melted_cormat <- reshape2::melt(upper_tri, na.rm = TRUE)

c2 <- ggplot(data = melted_cormat, aes(Var2, Var1, fill = value)) +
 geom_tile(color = "white")+
 scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
   midpoint = 0, limit = c(-1,1), space = "Lab", 
   name="Pearson\nCorrelation") +
  theme_minimal() + 
  theme(axis.text.x = element_text(angle = 45, vjust = 1, 
    size = 12, hjust = 1), 
    axis.text.y = element_text(size= 12))+
 coord_fixed()

c2 + 
  geom_text(aes(Var2, Var1, label = value), color = "black", size = 4) +
  theme(axis.title.x = element_blank(),
  axis.title.y = element_blank(),
  panel.grid.major = element_blank(),
  panel.border = element_blank(),
  panel.background = element_blank(),
  axis.ticks = element_blank(),
  legend.justification = c(1, 0),
  legend.position = c(0.6, 0.7),
  legend.direction = "horizontal")+
  guides(fill = guide_colorbar(barwidth = 7, barheight = 1,
                title.position = "top", title.hjust = 0.5)) +
  scale_x_discrete(labels = c("PDO","Harvest", "Recruits", "Upwelling", "SRKW", "Hatcheries", "Year", "CSL")) +
  scale_y_discrete(labels = c("PDO","Harvest", "Recruits", "Upwelling", "SRKW", "Hatcheries", "Year", "CSL"))

```

#### - Strong positive correlations: Year X CSL, Hatcheries X CSL, Hatcheries X SRKW (!) 
#### - Strong negative correlation: PDO X CSL, PDO X Recruits
#### - Weak but notable: Recruits are positively linked to CSL but negative to SRKW


### Time to follow up simple correlations with some CCM results ...

