---
title: "3_EDM_CCM_pairwise_causality"
author: "Alison Iles"
date: "6/27/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#load the necessary packages. 
Note that the `echo = FALSE` parameter prevents printing of the R code.

```{r, echo=FALSE}
library(rEDM)
library(reshape2)
library(ggplot2)
library(viridis)
library(gridExtra)
library(xtable)
library(rlist)
library(Kendall) #for MannKendall test
library(psych) #for paired.r function
library(tidyr)
library(stringr)
```


#Convergent Cross Mapping (CCM) with time delays
To test whether pairwise causality exists between the environmental variables and the spawner recruit relationship. Also to test all the lags for each environmental variable to figure out which is the best to use in the multivariate embeddings.

block_lnlp uses multiple time series given as input to generate an attractor reconstruction, and then applies the simplex projection or s-map algorithm to make forecasts. This method generalizes the simplex and s_map routines, and allows for "mixed" embeddings, where multiple time series can be used as different dimensions of an attractor reconstruction.

#Multivariate CCM for the whole ESU
We use the block_lnlp function to perform the cross-mappings in order to include effective spawners in the attractor reconstruction

```{r}
###--- nonlinear test for different major population groups
rm(list=ls())  
load("Data/Rdata/block_data.Rdata")  
      aa <- t(data.frame(as.list(names(block_data[[1]]))))
      rownames(aa) <- NULL #remove rownames
      bb <- data.frame(c(aa[13:nrow(aa)] ))
      colnames(bb) <- c("variable")
      cc <- str_split_fixed(bb$variable, "[.]",n=3)
var_names <- cbind(bb,cc)
      colnames(var_names) <- c("name","cat","subcat","offset")

data <- block_data #Include all mpgs at the whole ESU level
esuname <- as.character("SnakeESU")
  
    #concatenate the time series of each variable together
    keys <-c(names(data[[1]]))
    merged_data <-setNames(do.call(mapply, c(FUN=c, lapply(data, `[`, keys))), keys)
    merged_data <- as.data.frame(merged_data)
    valid <-  is.finite(merged_data$salm.rec3_n.0) & is.finite(merged_data$salm.rec4_n.0) & is.finite(merged_data$salm.rec5_n.0) & is.finite(merged_data$salm.eff_n.0)
    block <- merged_data[valid,]  
    
    #List in lib the begin and end points of each stock and the break points within stocks
    lib <- matrix(NA, nrow = length(which(diff(block$year)!=1))+1, ncol = 2)
    lib[,1] <- c(1, which(diff(block$year)!=1)+1)
    lib[,2] <- c(which(diff(block$year)!=1), nrow(block))
    minlib <- lib[,2]-lib[,1]
    lib <- lib[minlib>20,] #only include in the library the sections of data that are continuous for at least 20 time points. 
    x <- block[lib[1,1]:lib[1,2],] #narrow the data block to these library sections
        for (r in 2:nrow(lib)){
            xtmp <- block[lib[r,1]:lib[r,2],]
            x <- rbind(x,xtmp)
        }
   block <- x
   # need to recreate the library list
   lib <- matrix(NA, nrow = length(which(diff(block$year)!=1))+1, ncol = 2)
            lib[,1] <- c(1, which(diff(block$year)!=1)+1)
            lib[,2] <- c(which(diff(block$year)!=1), nrow(block))
        rm(x,minlib, xtmp, r)
    
    
    libsize <- c(50,100,150,300,450,700,1000,NROW(block))  
    
    #set up output table
    library_vars <- c("salm.rec3_n.0", "salm.rec4_n.0", "salm.rec5_n.0", "salm.rec_n.0")
    
    #Determine best embedding dimension: E would be set to the best univariate embedding dimension for each lib_column variable:
    best_E <- matrix(NA, nrow = length(library_vars), ncol = 2) # Save the optimal embedding dimension for each library variable
    colnames(best_E) <- c("library","E")
    best_E[,1] <- t(library_vars)
    for(j in 1:length(library_vars)){
      data <- as.matrix(cbind(block$year, block[library_vars[j]]))
      rownames(data) <- NULL #remove rownames to supress error in CCM: NAs introduced by coersion
            simplex_output <- simplex(data, E = 1:10, silent = TRUE)
            best_E[j,2] <-  min(simplex_output$E[simplex_output$rho>max(simplex_output$rho)-sd(simplex_output$rho)]) #choose the smallest E whose rho is within 1 stdv of the maximum rho
    }

    env_vars <- as.character(var_names[,1])
    CCM_outputlist <- expand.grid(env_vars, library_vars, stringsAsFactors = FALSE) 
    CCM_outputlist$temp <- CCM_outputlist$Var1
    CCM_outputlist <- separate(CCM_outputlist,temp, into=c("cat","subcat","offset"), sep = "[.]")           
    CCM_outputlist[,6:12] <- matrix(NA, nrow = NROW(CCM_outputlist), ncol = 7)
    colnames( CCM_outputlist) <- c("target","library","cat","subcat","offset","rho", "rho_eff","N", "95p_crit", "ccm_rho", "MannKendall", "FisherZ")
    CCM_outputlist <- merge(CCM_outputlist, best_E)
    

    for(i in c(1:NROW(CCM_outputlist)) ){  
    
      library_var <- CCM_outputlist$library[i]
      target_var <- CCM_outputlist$target[i]
      
            #make lagged data block that matches the optimal E
            max_lag <- as.numeric(as.character(CCM_outputlist$E[i]))
                  if (max_lag==0){max_lag=1}
            library_block <- make_block(block[library_var], t=block$year, max_lag=max_lag, tau=1, lib=lib)
            CCMlibrarynames <- names(library_block)
            CCMlibrarynames <- CCMlibrarynames[-c(1)] #remove "time" from list of library column names with which to do the CCM
            #CCMlibrarynames <- as.character(cbind(t(CCMlibrarynames),"block$eff_n") )

#---------------------------------------------------------------------------------------------------------  
# Univariate CCM with block_lnlp function 
            target_block <- make_block(block[target_var],t=block$year, max_lag=2, tau=1, lib=lib)
            lnlp_block <- as.matrix(cbind(library_block, target_block[target_var]))
            rownames(lnlp_block) <- NULL #remove rownames to supress error in CCM: NAs introduced by coersion
      
xmap <-  block_lnlp(lnlp_block, lib = lib, columns = CCMlibrarynames, target_column = target_var, method = c("simplex"), tp = 0, first_column_time = TRUE, silent = TRUE)
            CCM_outputlist$rho[i] <- xmap$rho
            CCM_outputlist$N[i] <- xmap$num_pred
            CCM_outputlist$`95p_crit`[i] <- qnorm(0.95, sd = 1/sqrt(xmap$num_pred - 3)) #qnorm is used to look up percentiles of the standard normal distribution. The 0.95 quantile is the 95th percentile. qnorm produces the boundary value that the rho needs to be greater than in order to be signifcantly different from zero and p=0.05 level. 

            
#---------------------------------------------------------------------------------------------------------  
# Multivariate CCM with block_lnlp function, including the number of effective spawners in the embedding 
            eff_data <- make_block(block["salm.eff_n.0"],t=block$year, max_lag=2, tau=1, lib=lib)
            lnlp_block_eff <- as.matrix(cbind(library_block[,1:dim(library_block)[2]-1], eff_data["salm.eff_n.0"], target_block[target_var]))
            rownames(lnlp_block_eff) <- NULL #remove rownames to supress error in CCM: NAs introduced by coersion
            CCMlibrarynames <- as.character(cbind(t(CCMlibrarynames),"salm.eff_n.0") )
      
xmap_eff <-  block_lnlp(lnlp_block_eff, lib = lib, columns = CCMlibrarynames, target_column = target_var, method = c("simplex"), tp = 0, first_column_time = TRUE, silent = TRUE)
             CCM_outputlist$rho_eff[i] <- xmap_eff$rho

#---------------------------------------------------------------------------------------------------------  
# Univariate CCM with CCM function, includes calculation of convergence 
            ccm_block <- as.matrix(cbind(library_block[,1:2],target_block[target_var]))
            rownames(ccm_block) <- NULL #remove rownames to supress error in CCM: NAs introduced by coersion
ccm <-  ccm(ccm_block, lib = lib, E = as.numeric(as.character(CCM_outputlist$E[i])), lib_column = library_var, target_column = target_var, first_column_time = TRUE, tp = 0, lib_sizes = libsize, random_libs = TRUE, num_samples=100, replace=TRUE, RNGseed=2301, silent = TRUE)      
      
      # Test if rho is significantly different from 0 at p=0.05 level (done for Sockey in Ye et al. 2015)
      # Calculate the median, maximum, and 1st & 3rd quantile of rho
            rho_quant=as.matrix(aggregate(ccm[,c('rho')],by = list(as.factor(ccm$lib_size)), quantile, na.rm = TRUE)[,'x'])
            rho_quant <- rho_quant[complete.cases(rho_quant*0),1:5] #removes rows with inf values
      CCM_outputlist$ccm_rho[i] <- signif(rho_quant[dim(rho_quant)[1],3],2) #save the cross map skill at max library size (50th percentile)
            
      #Test for monotonic trend in a time series z[t] based on the Kendall rank correlation of z[t] and t. 
            #Here the median, maximum, and 1st & 3rd quantiles of rho are tested, all need to be <0.05.
            ccmq=as.matrix(aggregate(ccm[,c('rho')],by = list(as.factor(ccm$lib_size)), quantile)[,'x'])
            ccmMK <- apply(ccmq[,2:5],2,MannKendall)
      CCM_outputlist$MannKendall[i] <- (ccmMK$`25%`[[2]]<0.05 & ccmMK$`50%`[[2]]<0.05 & ccmMK$`75%`[[2]]<0.05) # & ccmMK$`100%`[[2]]<0.05) not the last quantile as the plot often flattens out       
      # Tests that the max library z is significantly higher than the begining. 
      # Transforms the rho at the min library and max library length to a normally distributed Fisher's z value
            # Independent correlations, different sample sizes. Our small sample sizes are a disadvantage here.
      FZ <- paired.r(rho_quant[1,3],rho_quant[dim(rho_quant)[1],3], NULL, libsize[1], libsize[dim(rho_quant)[1]], twotailed=FALSE)
            CCM_outputlist$FisherZ[i] <- FZ$p            
#-----------------------------------------------------------------------------------------------------  
    }
    
saveRDS(CCM_outputlist, file = paste("Output/Rdata/3_CCM/3_CCM_",esuname,".RDS", sep = ""), compress = FALSE)


```
#CCM output list meta data:
    "library": In CCM analysis the 'library' variable is the variable of interest, in our case chinook recruitment
    "target": In CCM analysis the 'target' variable is the putative causal variable, in our case the various abiotic and biotic variables that are hypothesized to affect chinook recruitment
    "cat": The broader category that the target variable is in
    "subcat": The specific variable in the broader target variable categories
    "offset": The number of years that the target variable is offset. Environmental data are offset from 1 to 5 years after brood year. Pinniped data are offset from 2 to 5 years after brood year - assuming seals and sea lions eat smolts and returns. For the CSL pup count data, 6 years are added to each offset as only 6+ year old sea lion males swim up to Oregon and eat salmon (ref in Chasco paper). Orca data is offset by 4 to 5 years after brood year - assuming orcas just eat the returning chinook. 

The first analysis uses the block_lnlp function on a delay coordinate embedding of the library variable to predict the target variable. The embedding dimension used is the optimal E for the library variable (our variable of interest). 
    "rho": The correlation coefficient between observations and predictions from the block_lnlp function. 
    "rho_eff": The correlation coefficient from a parallel analysis that includes the number of effective spawners in place of one of the delay embeddings.  
    "N": The number of predictions made in the block_lnlp analysis
    "95p_crit": Based on the standard normal distribution, this is the boundary value that the rho needs to be greater than for a 95% confidence that rho is greater than zero. 

The second analysis uses the ccm function to test the univariate causal influence of the target variable on the library variable. The analysis tests many different library sizes sampled randomly many times to get distributions of rho values (the correlation coefficient between observations and predictions).
    "ccm_rho": The cross map skill at the maximum library size. 
    "MannKendall": A test for convergence, or an increasing monotonic trend in rho as library size increases based on the Kendall rank correlation. Here the first, median and & 3rd quantiles of rho are tested, not the last quantile as the plot often flattens out; all need to be <0.05 for this to be 'TRUE'.  
    "FisherZ": A test that rho at the max library size is significantly higher than at the minimum. 
    
“E”: the embedding dimension
“mpg”: the MPG or ESU region
“drho”; the delta rho representing the change and rho going from the univariate model to the multivariate model. Positive means the multivariate model had a higher rho and including effective spawners in the embedding improved the forecasting of the putative causal variable. 
“regionlabels”: a factor version of the mpg variable for plotting purposes
“salmoncohortlabels”: a factor version of the salmon recruit library variable for plotting purposes
“catlabels”: a factor version of the target variable categories for plotting purposes 
```{R}
#Figures
rm(list=ls())  
load("Data/Rdata/block_data.Rdata")  
  
library(ggplot2)

mpg <- c() 
  for(stk in 1:length(block_data)) { 
      mpg[stk] <- block_data[[stk]]$mpg[1]
  } 

for(u in c(0,1,3,5)){
  if(u==0){ 
    mpgname <- "Snake River ESU"
    CCM_outputlist<- readRDS("Output/Rdata/3_CCM/3_CCM_SnakeESU.RDS" )}
  if(u>0){
  data <- block_data[c(mpg==u)]
  mpgname <- as.character(data[[1]]$mpg[1])
  CCM_outputlist<- readRDS(paste("Output/Rdata/3_CCM/3_CCM_MPG_",mpgname,".RDS", sep = ""))}

    # sort output table by offset then category
    d <-  CCM_outputlist[order(CCM_outputlist$offset),]
    d <-  d[order(d$cat),]
    d <-  d[order(d$subcat),]
    d$offset <- as.numeric(d$offset)
    d$mpg <- mpgname
 
    if(u==0){D <- d}
    if(u>0){D <- rbind(D,d)}
}

D$drho <- D$rho_eff - D$rho

D$regionlabels <- factor(D$mpg, levels=c("Snake River ESU", "Imnaha", "Middle Fork Salmon", "Upper Salmon"), labels=c("Snake River ESU", "Imnaha MPG", "Middle Fork Salmon MPG", "Upper Salmon MPG"))

D$salmoncohortlabels <- factor(D$library, levels=c("salm.rec_n.0","salm.rec3_n.0", "salm.rec4_n.0", "salm.rec5_n.0"), labels=c("All recruits", "3 yr old recruits", "4 yr old recruits", "5 yr old recruits"))

manual_color_codes <- read.csv("Data/csv_data/CORE_CCM_figure_color_codes.csv")

varcat <- unique(D$cat)
D$catlabels <- factor(D$cat, levels=c("hatch", "npgo", "pdo", "upw", "csl", "ssl", "harv", "hseal", "salm", "arc", "orca", "flow" ), labels=c("Hatcheries", "NPGO", "PDO", "Upwelling", "California sea lions", "Steller sea lions", "Chinook harvest", "Harbor seals", "Chinook salmon", "Sea surface temperature", "Orca whales", "River flow"))

saveRDS(D, file = "Output/Rdata/3_CCM/3_CCM_figure_data.RDS", compress = FALSE)



#Multivariate
for(c in c(1:length(varcat))) {
  D_plot <- D[D$cat==varcat[c],]
  mcc <- manual_color_codes[manual_color_codes$cat==varcat[c],2:4]
  mcc <- sapply(mcc, unlist)
  rownames(mcc) <- mcc[,2]
  
  p1 <- ggplot(D_plot, aes(x=offset, y=rho_eff)) +
    geom_line(aes(col=subcat)) +
    geom_point(aes(col=subcat), show.legend = TRUE) + 
    geom_vline(aes(xintercept=0), colour='#999999') +
    geom_hline(aes(yintercept=0), colour='#999999') +
    geom_hline(aes(yintercept=min(D_plot$'95p_crit')), colour='#999999', linetype="dashed") + #horizontal line for rho sig diff from zero
    theme_bw() + 
    guides(col = guide_legend(ncol=1)) +
    labs(title= D_plot$catlabels[1], subtitle = "Multivariate CCM: uses the block_lnlp function; includes effective spawners in the embedding ", x="Variable offset, year", y=expression(paste("Cross-mapping correlation, ", {rho}))) + 
    scale_color_manual(values = mcc[,1], name="Time series", breaks = mcc[,2], labels = mcc[,3]) 
    
p2 <- p1 + facet_grid(regionlabels ~ salmoncohortlabels) 

ggsave(filename = paste("Output/Figures/3_CCM/3_CCM_",varcat[c],"_multivariate.pdf", sep = ""), plot = p2, width = 7, height = 9.5, units = "in")  
}

#Univariate
for(c in c(1:length(varcat))) {
  D_plot <- D[D$cat==varcat[c],]
  mcc <- manual_color_codes[manual_color_codes$cat==varcat[c],2:4]
  mcc <- sapply(mcc, unlist)
  rownames(mcc) <- mcc[,2]
  
  p1 <- ggplot(D_plot, aes(x=offset, y=rho)) +
    geom_line(aes(col=subcat)) +
    geom_point(aes(col=subcat, shape=(MannKendall==TRUE), show.legend = TRUE)) +
    geom_vline(aes(xintercept=0), colour='#999999') +
    geom_hline(aes(yintercept=0), colour='#999999') +
    geom_hline(aes(yintercept=min(D_plot$'95p_crit')), colour='#999999', linetype="dashed") + #horizontal line for rho sig diff from zero
    theme_bw() + 
    guides(col = guide_legend(ncol=1)) +
    labs(title= D_plot$catlabels[1], subtitle = "Univariate CCM: uses the ccm function; without effective spawners in the embedding ", x="Variable offset, year", y=expression(paste("Cross-mapping correlation, ", {rho}))) + 
    scale_color_manual(values = mcc[,1], name="Time series", breaks = mcc[,2], labels = mcc[,3]) +
    scale_shape_manual(name="MK test",  values=c("FALSE"=1, "TRUE"=16),  breaks = c("TRUE", "FALSE"), labels = c("True", "False"))

p2 <- p1 + facet_grid(regionlabels ~ salmoncohortlabels) 

ggsave(filename = paste("Output/Figures/3_CCM/3_CCM_",varcat[c],"_univariate.pdf", sep = ""), plot = p2, width = 7, height = 9.5, units = "in")  
}


#change in rho from univariate to multivariate embeddings
for(c in c(1:length(varcat))) {
  D_plot <- D[D$cat==varcat[c],]
  mcc <- manual_color_codes[manual_color_codes$cat==varcat[c],2:4]
  mcc <- sapply(mcc, unlist)
  rownames(mcc) <- mcc[,2]
  
  p1 <- ggplot(D_plot, aes(x=offset, y=drho)) +
    geom_line(aes(col=subcat)) +
    geom_point(aes(col=subcat, show.legend = TRUE)) +
    geom_vline(aes(xintercept=0), colour='#999999') +
    geom_hline(aes(yintercept=0), colour='#999999') +
    geom_hline(aes(yintercept=min(D_plot$'95p_crit')), colour='#999999', linetype="dashed") + #horizontal line for rho sig diff from zero
    theme_bw() + 
    guides(col = guide_legend(ncol=1)) +
    labs(title= D_plot$catlabels[1], subtitle = "The change in rho from univariate pairwise CCM to multivariate CCM that includes the number of effective spawners in the embeddings ", x="Variable offset, year", y=expression(paste("Cross-mapping correlation, ", {rho}))) + 
    scale_color_manual(values = mcc[,1], name="Time series", breaks = mcc[,2], labels = mcc[,3])


p2 <- p1 + facet_grid(regionlabels ~ salmoncohortlabels) 

ggsave(filename = paste("Output/Figures/3_CCM/3_CCM_",varcat[c],"_delta_rho.pdf", sep = ""), plot = p2, width = 7, height = 9.5, units = "in")  
}
 
```