---
title: "4_EDM_Univariate_forecast_models"
author: "Alison Iles + Kurt Ingeman"
date: "6/27/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#load the necessary packages. 
Note that the `echo = FALSE` parameter prevents printing of the R code.

```{r, echo=FALSE}
library(rEDM)
#library(rjags)
library(reshape2)
library(rgl)
library(ggplot2)
library(gridExtra)
library(xtable)
library(rlist)
```

After CCM, the next step is to produce univariate forecast models. Recall that the best predictions were obtained using a high number of time lags - the dynamics were high dimensional. As such, it could be that adding additional information from other variables will improve predictions. Specifically, we compare models that predict recruits using lags of itself (univariate) vs. models that also include environmental variables (multivariate).

We can do this using the block_lnlp() function, but first we need to construct lags for each variable. Here, we use the make_block() function, which automatically adds lags of the variables. However, we also need to be careful, since the raw data combines observations from multiple populations. By including the lib argument, we can indicate which parts of the time series correspond to different segments, so that lags indicate unknown values correctly.


```{r}
load("Data/Rdata/block_data.Rdata")

#Preparing the data for one major population group
#construct a single composite time series with the normalized data

#ID the different mpg
    mpg <- c() 
      for(stk in 1:length(block_data)) { 
          mpg[stk] <- block_data[[stk]]$mpg[1]
      } 
Imnaha <- block_data[c(mpg==1)] #Only Imnaha
Imnaha <- rbind(Imnaha[[1]],Imnaha[[2]],Imnaha[[3]],Imnaha[[4]],Imnaha[[5]],Imnaha[[6]]) #merge data into one data frame
Imnaha <- Imnaha[c("stk", "yr", "eff_n", "rec3_n", "rec4_n", "rec5_n", "ret_n", "up_apr", "up_oct", "up_nov", "pdo_win")] #keep time, population, normalized spawners and rec4, environmental vars
Imnaha <- droplevels(Imnaha) #drop unused levels from factors in a data frame
# To prevent lagged vectors from being constructed that span separate populations, we need to create an appropriae index variable to identify different segments. We can then assess the predictive skill of EDM by using cross-validation and selecting random subsets of populations to use for the library and prediction sets
      data_by_popn <- split(Imnaha, Imnaha$stk)
      segments_end <- cumsum(sapply(data_by_popn, NROW))
      segments_begin <- unname(c(1, segments_end[-length(segments_end)] + 1))
      segments <- cbind(segments_begin, segments_end)
       # Choose random segments for prediction
      set.seed(7583)
      rndlib <- sample(1:NROW(segments), floor(NROW(segments) * 0.75))
      composite_lib <- segments[rndlib, ]  
      composite_pred <- segments[-rndlib, ]

      
block_Imnaha <- make_block(Imnaha[, 2:11], t = Imnaha$yr, max_lag = 5,
lib = segments) 

# Note that the lagged columns begin with NA values because there are no observations of the variables for times t < 1. The vectors that include NA values are excluded if that specific value is needed for computation.

```

Next, we run the models, being sure to set up predictor variables (the columns argument) and the variable to be predicted (the target_column argument). By default, predictions are always for one step ahead. Note that some models include only lagged observations of the target variable, and other models also include environmental data




```{r}
block_lnlp_4 <- function(block, target_column, columns, norm = FALSE)
{
    if(norm)
    {
        block[,columns] <- normalize(block[,columns])
    }
    
    lib_segments <- matrix(NA, nrow = 4, ncol = 2)
    segment_size <- NROW(block)/4
    start_index <- 1
    for(i in 1:4)
    {
        lib_segments[i,1] <- floor(start_index)
        end_index <- start_index - 1 + segment_size
        lib_segments[i,2] <- floor(end_index)
        start_index <- end_index+1
    }
    
    if(is.list(columns))
    {
        preds <- lapply(1:length(columns), function(x) {rep.int(NA, times = NROW(block))})
        for(i in 1:4)
        {
            pred_index <- lib_segments[i,1]:lib_segments[i,2]
            
            temp <- block_lnlp(block, lib = lib_segments[-i,], pred = lib_segments[i,], 
                               target_column = target_column, tp = 0, 
                               first_column_time = TRUE, 
                               columns = columns, stats_only = FALSE)
            
            for(j in 1:length(columns))
                preds[[j]][pred_index] <- temp$model_output[[j]]$pred[pred_index] #AI: changed from temp[[j]]$model_output$pred[pred_index]
        }
    }
    else
    {
        preds <- rep.int(NA, times = NROW(block))
        for(i in 1:4)
        {
            pred_index <- lib_segments[i,1]:lib_segments[i,2]
            
            temp <- block_lnlp(block, lib = lib_segments[-i,], pred = lib_segments[i,], 
                               target_column = target_column, tp = 0, 
                               first_column_time = TRUE, 
                               columns = columns, stats_only = FALSE)
            
            preds[pred_index] <- temp$model_output[[1]]$pred[pred_index] #AI: changed from temp[[1]]$model_output$pred[pred_index]

        }
    }
    return(preds)
}


simple_EDM <- function()
{
    forecast <- function(stock_df)
    {  
        make_forecasts <- function(block, mu_3, sigma_3, mu_4, sigma_4, mu_5, sigma_5)
        {
            rec3 <- block_lnlp_4(block, target_column = 2, columns = 1) #use function block_lnlp_4
            rec4 <- block_lnlp_4(block, target_column = 3, columns = 1)
            rec5 <- block_lnlp_4(block, target_column = 4, columns = 1)
            
            rec3 <- rec3*sigma_3 + mu_3
            rec4 <- rec4*sigma_4 + mu_4
            rec5 <- rec5*sigma_5 + mu_5
            return(rec3 + c(NA, rec4[1:(NROW(block)-1)]) + c(NA, NA, rec5[1:(NROW(block)-2)]))
        }
        
        # set up recruits and spawners
        valid <- is.finite(stock_df$rec) & is.finite(stock_df$eff) #exclude NAs? Should we??
        returns <- stock_df$ret[valid] #NAs left in here...
        years <- stock_df$yr[valid]
        spawners <- stock_df$eff_n[valid]
        recruits_3 <- stock_df$rec3_n[valid]
        mu_3 <- stock_df$rec3_mu[valid]
        sigma_3 <- stock_df$rec3_sigma[valid]
        recruits_4 <- stock_df$rec4_n[valid]
        mu_4 <- stock_df$rec4_mu[valid]
        sigma_4 <- stock_df$rec4_sigma[valid]
        recruits_5 <- stock_df$rec5_n[valid]
        mu_5 <- stock_df$rec5_mu[valid]
        sigma_5 <- stock_df$rec5_sigma[valid]
        
        # make block
        block <- data.frame(years = years, eff = spawners, 
                            rec3=recruits_3, rec4 = recruits_4, rec5 = recruits_5)
        
        if(length(returns) < 2) # check for enough data
            return(data.frame(year = NaN, obs = NaN, pred = NaN))
        
        forecasts <- make_forecasts(block,  mu_3, sigma_3, mu_4, sigma_4, mu_5, sigma_5)
        return(data.frame(year = years, obs = returns, pred = forecasts))
    }
    
    load("Data/Rdata/block_data.Rdata")
    
    # make forecasts for each stock
    results <- lapply(names(block_data), 
                      function(stk_name) {
                          cat("forecasting for ", stk_name, "... ", sep = "")
                          start_time <- proc.time()
                          output <- forecast(block_data[[stk_name]])
                          elapsed_time <- proc.time() - start_time
                          cat("(", elapsed_time[3], " sec.)\n", sep = "")
                          return(output)
                      })
    names(results) <- names(block_data)
    saveRDS(results, file = "Output/Rdata/results_simple_EDM.RDS")
    
    # compute stats
    stats <- do.call(rbind, lapply(results, function(stock_results) {
        compute_stats(stock_results$obs, stock_results$pred)
    }))
    stats$stk <- names(block_data)
    saveRDS(stats, file = "Output/Rdata/stats_simple_EDM.RDS")
    return()
}

simple_EDM()

```