---
title: "4b_EDM_Univariate_forecast_models"
author: "Kurt Ingeman"
date: "9/10/19"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#load the necessary packages. 
Note that the `echo = FALSE` parameter prevents printing of the R code.

```{r, echo=FALSE}
library(rEDM)
library(dplyr)
#library(rjags)
library(reshape2)
library(rgl)
library(ggplot2)
library(gridExtra)
library(xtable)
library(rlist)
```

After CCM, the next step is to produce univariate forecast models. Recall that the best predictions were obtained using a high number of time lags - the dynamics were high dimensional. As such, it could be that adding additional information from other variables will improve predictions. Specifically, we compare models that predict recruits using lags of itself (univariate) vs. models that also include spawners and eventually environmental variables (multivariate).

We can do this using the block_lnlp() function, but first we need to construct lags for each variable. I will do so following the Ye tutorial (a combination of manual lagging and canned rEDM functions) rather than the make_block function. We also need to be careful, since the raw data combines observations from multiple stocks within each MPG. By including the lib argument, we can indicate which parts of the time series correspond to different segments, so that lags indicate unknown values correctly.

```{r}
### Preparing the data for one major population group at a time

data <- read.csv("Data/csv_data/SRSS_cohort_prelim.csv")

# Sort the data by stk name to avoid mismatch with segments
data <-  data %>% 
  arrange(stk, yr)

normalize <- function(x, ...) {
  (x - mean(x, ...))/sd(x, ...)
}

# the vars that we want to use for "univariate" prediction are simply the recruit cohorts and the number of spawners. Later we will bring in the env variables to improve forecasts

# separate time column from data
vars <- c("eff", "rec3", "rec4", "rec5", "rec")
composite_ts <- data[, vars]

# normalize each time series within an individual stock
data_by_stk <- split(composite_ts, data$stk)
normalized_data <- lapply(data_by_stk, function(df) sapply(df, normalize))
composite_ts <- cbind(Year = data$yr, data.frame(do.call(rbind,normalized_data)))
# KI Note: Now we have the year and 5 variables with no lagged variables

# To prevent lagged vectors from being constructed that span separate stks, 
# we need to create an appropriate index variable that identifies different segments.

segments_end <- cumsum(sapply(data_by_stk, NROW))
segments_begin <- c(1, segments_end[-length(segments_end)] + 1)
segments <- cbind( segments_begin, segments_end) ## reversed by KI because it introduced an unnamed stock at begining and pushed Bear Valley segment down into next stock

# Choose random segments for prediction
set.seed(9837)
rndlib <- sample(1:NROW(segments), floor(NROW(segments) * 0.75))
composite_lib <- segments[rndlib, ]
composite_pred <- segments[-rndlib, ]

# alternative when using leave-one-out, use all for lib and pred
set.seed(9837)
rndlib <- sample(1:NROW(segments), floor(NROW(segments)))
full_lib <- segments[rndlib, ]
full_pred <- full_lib

######################################################################
### THESE ARE NOT RANDOM SEGMENTS; THEY ARE DIFFERENT STOCKS
### WE NEED TO SPLIT BY STOCK AND CHOOSE RANDOM SEGMENTS WITHIN INSTEAD
### I need to modify the code from the sockeye paper for knowing where segments are but stick to leave one out validation using the entire dataset
######################################################################

set.seed(2312)
rndlib <- sample(1:NROW(data), floor(NROW(data) * 0.75))
comp_lib <- data[rndlib, ]
comp_pred <- data[-rndlib, ]

# Now we have normalized the data, split data into segments by plot, 
# and created a series of chunks for training and prediction...

#################
## Run simplex ##
## and S-map ####
#################

simplex_out <- lapply(names(composite_ts)[2:6], function(var) {
simplex(composite_ts[, c("Year", var)], E = 2:8, lib = full_lib, pred = full_pred)
})

names(simplex_out) <- names(composite_ts)[-1]

par(mar = c(4, 4, 1, 1), mfrow = c(2, 3), mgp = c(2.5, 1, 0))  # set up margins for plotting
out <- lapply(names(simplex_out), function(var) {
  plot(simplex_out[[var]]$E, simplex_out[[var]]$rho, type = "l", xlab = "Embedding Dimension (E)", 
       ylab = "Forecast Skill (rho)", main = var)
})

best_E <- sapply(simplex_out, function(df) {df$E[which.max(df$rho)]})
best_E

smap_out <- lapply(names(composite_ts)[2:6], function(var) {
  s_map(composite_ts[, c("Year", var)], E = best_E[var], lib = full_lib, 
        pred = full_pred)
})

names(smap_out) <- names(simplex_out)

par(mar = c(4, 4, 1, 1), mfrow = c(2, 3), mgp = c(2.5, 1, 0))  
# set up margins for plotting

lapply(names(smap_out), function(var) {
  plot(smap_out[[var]]$theta, smap_out[[var]]$rho, type = "l", xlab = "Nonlinearity (theta)", 
       ylab = "Forecast Skill (rho)", main = var)
})


################## 
## Multivariate ##
##### Models #####
################## 

data_by_stk <- split(composite_ts, data$stk)

lagged_data <- do.call(rbind, lapply(data_by_stk, function(df) {
  n <- NROW(df)
  temp <- data.frame(Year = df$Year)
  temp$rec_t <- df$rec
  temp$rec_tm1 <- c(NA, temp$rec_t[-n])
  temp$rec_tm2 <- c(NA, temp$rec_tm1[-n])
  temp$rec_tm3 <- c(NA, temp$rec_tm2[-n])
  
  temp$eff_t <- df$eff
  temp$eff_tm1 <- c(NA, temp$eff_t[-n])
  temp$eff_tm2 <- c(NA, temp$eff_tm1[-n])
  temp$eff_tm3 <- c(NA, temp$eff_tm2[-n])
  
  temp$rec3_t <- df$rec3
  temp$rec3_tm1 <- c(NA, temp$rec3_t[-n])
  temp$rec3_tm2 <- c(NA, temp$rec3_tm1[-n])
  temp$rec3_tm3 <- c(NA, temp$rec3_tm2[-n])
  
  temp$rec4_t <- df$rec4
  temp$rec4_tm1 <- c(NA, temp$rec4_t[-n])
  temp$rec4_tm2 <- c(NA, temp$rec4_tm1[-n])
  temp$rec4_tm3 <- c(NA, temp$rec4_tm2[-n])
  
  temp$rec5_t <- df$rec5
  temp$rec5_tm1 <- c(NA, temp$rec5_t[-n])
  temp$rec5_tm2 <- c(NA, temp$rec5_tm1[-n])
  temp$rec5_tm3 <- c(NA, temp$rec5_tm2[-n])
  
  return(temp)
}))

head(lagged_data)


```


Run univariate and multivariate forecasts using block_lnlp function and plot the observed vs predicted

```{r}
univ_cols <- c("rec_t", "rec_tm1", "rec_tm2", "rec_tm3")
univ_output <- block_lnlp(lagged_data, lib = composite_lib, pred = composite_pred, 
                          columns = univ_cols , target_column = 1, stats_only = FALSE, first_column_time = TRUE)

univ_model_output <- univ_output[[1]]$model_output 

univ_observed <- univ_model_output$obs
univ_predicted <- univ_model_output$pred

plot_range <- range(c(univ_observed, univ_predicted), na.rm = TRUE)
plot(univ_observed, univ_predicted, xlim = plot_range, ylim = plot_range, xlab = "Observed", 
     ylab = "Predicted", asp = 1)
abline(a = 0, b = 1, lty = 2, col = "blue")

multi_cols <- c("rec_t", "rec_tm1", "rec_tm2", "rec_tm3", "eff_t", "eff_tm1","eff_tm2","eff_tm3")
multi_output <- block_lnlp(lagged_data, lib = composite_lib, pred = composite_pred, columns = multi_cols , target_column = 1, stats_only = FALSE, first_column_time = TRUE)

multi_model_output <- multi_output[[1]]$model_output 

multi_observed <- multi_model_output$obs
multi_predicted <- multi_model_output$pred


par(mar = c(4, 4, 1, 1), mgp = c(2.5, 1, 0), pty = "s")  # set up margins for plotting
plot_range <- range(c(univ_observed, univ_predicted), na.rm = TRUE)
plot(univ_observed, univ_predicted, xlim = plot_range, ylim = plot_range, xlab = "Observed", 
     ylab = "Predicted")
abline(a = 0, b = 1, lty = 2, col = "darkgrey", lwd = 2)
abline(lm(univ_predicted ~ univ_observed), col = "black", lty = 3, lwd = 2)

points(multi_observed, multi_predicted, pch = 2, col = "red")
abline(lm(multi_predicted ~ multi_observed), col = "red", lty = 3, lwd = 2)

legend("topleft", legend = c(paste("rho =", round(univ_output[[1]]$stats$rho, 
                                                  2)), paste("rho =", round(multi_output[[1]]$stats$rho, 2))), lty = 3, lwd = 2, 
       col = c("black", "red"), bty = "n")
```
Now, let's split recruits by cohort such that forecasts can be assembled into return estimates. Here looking at whether eff spawners improves predictions for rec 4

```{r}
data_by_stk <- split(composite_ts, data$stk)

lag_rec4 <- do.call(rbind, lapply(data_by_stk, function(df) {
  n <- NROW(df)
  temp <- data.frame(Year = df$Year)
  
  temp$rec3_t <- df$rec3
  temp$rec3_tm1 <- c(NA, temp$rec3_t[-n])
 
  temp$rec4_t <- df$rec4
  temp$rec4_tm1 <- c(NA, temp$rec4_t[-n])
  temp$rec4_tm2 <- c(NA, temp$rec4_tm1[-n])
  temp$rec4_tm3 <- c(NA, temp$rec4_tm2[-n])
  temp$rec4_tm4 <- c(NA, temp$rec4_tm3[-n]) 
  temp$rec4_tm5 <- c(NA, temp$rec4_tm4[-n])
  
  temp$eff_t <- df$eff
  temp$eff_tm1 <- c(NA, temp$eff_t[-n])
  temp$eff_tm2 <- c(NA, temp$eff_tm1[-n])
  temp$eff_tm3 <- c(NA, temp$eff_tm2[-n])
  temp$eff_tm4 <- c(NA, temp$eff_tm3[-n])
  temp$eff_tm5 <- c(NA, temp$eff_tm4[-n])
  
  
  return(temp)
}))

univ_cols <- c("rec4_t", "rec4_tm1", "rec4_tm2", "rec4_tm3", "rec4_tm4")
univ_output <- block_lnlp(lag_rec4, lib = full_lib, pred = full_pred,columns = univ_cols, target_column = 1, stats_only = FALSE, first_column_time = TRUE)

univ_model_output <- univ_output[[1]]$model_output 

univ_observed <- univ_model_output$obs
univ_predicted <- univ_model_output$pred

univ_output[[1]]$stats 


#### And multivariate, using eff and lags of eff

multi_cols <- c("rec4_t", "rec4_tm1", "rec4_tm2", "rec3_t", "eff_t")
multi_output <- block_lnlp(lag_rec4, lib = full_lib, pred = full_pred, columns = multi_cols , target_column = 1, stats_only = FALSE, first_column_time = TRUE)

multi_model_output <- multi_output[[1]]$model_output 

multi_observed <- multi_model_output$obs
multi_predicted <- multi_model_output$pred

multi_output[[1]]$stats
```

data_by_stk <- split(composite_ts, data$stk)

lag_rec5 <- do.call(rbind, lapply(data_by_stk, function(df) {
  n <- NROW(df)
  temp <- data.frame(Year = df$Year)
  temp$rec3_t <- df$rec3
  temp$rec3_tm1 <- c(NA, temp$rec3_t[-n])
 
  temp$rec4_t <- df$rec4
  temp$rec4_tm1 <- c(NA, temp$rec4_t[-n])
  temp$rec4_tm2 <- c(NA, temp$rec4_tm1[-n])
  temp$rec4_tm3 <- c(NA, temp$rec4_tm2[-n])
  temp$rec4_tm4 <- c(NA, temp$rec4_tm3[-n]) 
  temp$rec4_tm5 <- c(NA, temp$rec4_tm4[-n])
  
  temp$rec5_t <- df$rec5
  temp$rec5_tm1 <- c(NA, temp$rec5_t[-n])
  temp$rec5_tm2 <- c(NA, temp$rec5_tm1[-n])
  temp$rec5_tm3 <- c(NA, temp$rec5_tm2[-n])
  temp$rec5_tm4 <- c(NA, temp$rec5_tm3[-n]) 
  temp$rec5_tm5 <- c(NA, temp$rec4_tm4[-n])
  
  temp$eff_t <- df$eff
  temp$eff_tm1 <- c(NA, temp$eff_t[-n])
  temp$eff_tm2 <- c(NA, temp$eff_tm1[-n])
  temp$eff_tm3 <- c(NA, temp$eff_tm2[-n])
  temp$eff_tm4 <- c(NA, temp$eff_tm3[-n])
  temp$eff_tm5 <- c(NA, temp$eff_tm4[-n])
  
  
  return(temp)
}))

univ_cols <- c("rec5_t", "rec5_tm1", "rec5_tm2", "rec5_tm3", "rec5_tm4", "rec5_tm5")

univ_output <- block_lnlp(lag_rec4, lib = full_lib, pred = full_pred,columns = univ_cols, target_column = 1, stats_only = FALSE, first_column_time = TRUE)

univ_model_output <- univ_output[[1]]$model_output 

univ_observed <- univ_model_output$obs
univ_predicted <- univ_model_output$pred

#### And multivariate, using eff and lags of eff

multi_cols <- c("rec4_t", "rec4_tm1", "rec4_tm2", "rec4_tm3", "rec3_t", "eff_t")
multi_output <- block_lnlp(lag_rec4, lib = full_lib, pred = full_pred, columns = multi_cols , target_column = 1, stats_only = FALSE, first_column_time = TRUE)

multi_model_output <- multi_output[[1]]$model_output 

multi_observed <- multi_model_output$obs
multi_predicted <- multi_model_output$pred

multi_output[[1]]$stats
```



Next, we run the models, being sure to set up predictor variables (the columns argument) and the variable to be predicted (the target_column argument). By default, predictions are always for one step ahead. Note that some models include only lagged observations of the target variable, and other models also include environmental data




```{r}
block_lnlp_4 <- function(block, target_column, columns, norm = FALSE)
{
    if(norm)
    {
        block[,columns] <- normalize(block[,columns])
    }
    
    lib_segments <- matrix(NA, nrow = 4, ncol = 2)
    segment_size <- NROW(block)/4
    start_index <- 1
    for(i in 1:4)
    {
        lib_segments[i,1] <- floor(start_index)
        end_index <- start_index - 1 + segment_size
        lib_segments[i,2] <- floor(end_index)
        start_index <- end_index+1
    }
    
    if(is.list(columns))
    {
        preds <- lapply(1:length(columns), function(x) {rep.int(NA, times = NROW(block))})
        for(i in 1:4)
        {
            pred_index <- lib_segments[i,1]:lib_segments[i,2]
            
            temp <- block_lnlp(block, lib = lib_segments[-i,], pred = lib_segments[i,], 
                               target_column = target_column, tp = 0, 
                               first_column_time = TRUE, 
                               columns = columns, stats_only = FALSE)
            
            for(j in 1:length(columns))
                preds[[j]][pred_index] <- temp$model_output[[j]]$pred[pred_index] #AI: changed from temp[[j]]$model_output$pred[pred_index]
        }
    }
    else
    {
        preds <- rep.int(NA, times = NROW(block))
        for(i in 1:4)
        {
            pred_index <- lib_segments[i,1]:lib_segments[i,2]
            
            temp <- block_lnlp(block, lib = lib_segments[-i,], pred = lib_segments[i,], 
                               target_column = target_column, tp = 0, 
                               first_column_time = TRUE, 
                               columns = columns, stats_only = FALSE)
            
            preds[pred_index] <- temp$model_output[[1]]$pred[pred_index] #AI: changed from temp[[1]]$model_output$pred[pred_index]

        }
    }
    return(preds)
}


simple_EDM <- function()
{
    forecast <- function(stock_df)
    {  
        make_forecasts <- function(block, mu_3, sigma_3, mu_4, sigma_4, mu_5, sigma_5)
        {
            rec3 <- block_lnlp_4(block, target_column = 2, columns = 1) #use function block_lnlp_4
            rec4 <- block_lnlp_4(block, target_column = 3, columns = 1)
            rec5 <- block_lnlp_4(block, target_column = 4, columns = 1)
            
            rec3 <- rec3*sigma_3 + mu_3
            rec4 <- rec4*sigma_4 + mu_4
            rec5 <- rec5*sigma_5 + mu_5
            return(rec3 + c(NA, rec4[1:(NROW(block)-1)]) + c(NA, NA, rec5[1:(NROW(block)-2)]))
        }
        
        # set up recruits and spawners
        valid <- is.finite(stock_df$rec) & is.finite(stock_df$eff) #exclude NAs? Should we??
        returns <- stock_df$ret[valid] #NAs left in here...
        years <- stock_df$yr[valid]
        spawners <- stock_df$eff_n[valid]
        recruits_3 <- stock_df$rec3_n[valid]
        mu_3 <- stock_df$rec3_mu[valid]
        sigma_3 <- stock_df$rec3_sigma[valid]
        recruits_4 <- stock_df$rec4_n[valid]
        mu_4 <- stock_df$rec4_mu[valid]
        sigma_4 <- stock_df$rec4_sigma[valid]
        recruits_5 <- stock_df$rec5_n[valid]
        mu_5 <- stock_df$rec5_mu[valid]
        sigma_5 <- stock_df$rec5_sigma[valid]
        
        # make block
        block <- data.frame(years = years, eff = spawners, 
                            rec3=recruits_3, rec4 = recruits_4, rec5 = recruits_5)
        
        if(length(returns) < 2) # check for enough data
            return(data.frame(year = NaN, obs = NaN, pred = NaN))
        
        forecasts <- make_forecasts(block,  mu_3, sigma_3, mu_4, sigma_4, mu_5, sigma_5)
        return(data.frame(year = years, obs = returns, pred = forecasts))
    }
    
    load("Data/Rdata/block_data.Rdata")
    
    # make forecasts for each stock
    results <- lapply(names(block_data), 
                      function(stk_name) {
                          cat("forecasting for ", stk_name, "... ", sep = "")
                          start_time <- proc.time()
                          output <- forecast(block_data[[stk_name]])
                          elapsed_time <- proc.time() - start_time
                          cat("(", elapsed_time[3], " sec.)\n", sep = "")
                          return(output)
                      })
    names(results) <- names(block_data)
    saveRDS(results, file = "Output/Rdata/results_simple_EDM.RDS")
    
    # compute stats
    stats <- do.call(rbind, lapply(results, function(stock_results) {
        compute_stats(stock_results$obs, stock_results$pred)
    }))
    stats$stk <- names(block_data)
    saveRDS(stats, file = "Output/Rdata/stats_simple_EDM.RDS")
    return()
}

simple_EDM()

```