---
title: "4_EDM_Univariate_forecast_models"
author: "Alison Iles"
date: "6/27/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#load the necessary packages. 
Note that the `echo = FALSE` parameter prevents printing of the R code.

```{r, echo=FALSE}
library(rEDM)
#library(rjags)
library(reshape2)
library(rgl)
library(ggplot2)
library(gridExtra)
library(xtable)
library(rlist)
```

After CCM, the next step is to produce univariate forecast models. Recall that the best predictions were obtained using a high number of time lags - the dynamics were high dimensional. As such, it could be that adding additional information from other variables will improve predictions. Specifically, we compare models that predict recruits using lags of itself (univariate) vs. models that also include environmental variables (multivariate).

We can do this using the block_lnlp() function, but first we need to construct lags for each variable. Here, we use the make_block() function, which automatically adds lags of the variables. However, we also need to be careful, since the raw data combines observations from multiple populations. By including the lib argument, we can indicate which parts of the time series correspond to different segments, so that lags indicate unknown values correctly.


```{r}
load("Data/Rdata/block_data.Rdata")

# Preparing the data for one major population group
# construct a single composite time series with the normalized data

# ID the different mpg
    mpg <- c() 
      for(stk in 1:length(block_data)) { 
          mpg[stk] <- block_data[[stk]]$mpg[1]
      } 
    
Imnaha <- block_data[c(mpg==1)] # Only Imnaha
Imnaha <- rbind(Imnaha[[1]],Imnaha[[2]],Imnaha[[3]],Imnaha[[4]],Imnaha[[5]],Imnaha[[6]]) # Merge lists of stocks into one dataframe
Imnaha <- Imnaha[c("stk", "yr", "eff_n", "rec3_n", "rec4_n", "rec5_n", "ret_n", "up48_apr", "up48_oct", "up48_nov", "pdo_win")] 
#keep time, population, normalized spawners and rec4, environmental vars
Imnaha <- droplevels(Imnaha) #drop unused levels from factors in a data frame

# To prevent lagged vectors from being constructed that span separate populations, we need to create an appropriae index variable to identify different segments. But, here we depart from Ye by employing leave-one-out forecasting on all sections. We still feed the function a matrix of beginning and end points for each stock, but now we pred and lib on all points, not a random selection stocks

      data_by_popn <- split(Imnaha, Imnaha$stk)
      segments_end <- cumsum(sapply(data_by_popn, NROW))
      segments_begin <- unname(c(1, segments_end[-length(segments_end)] + 1))
      segments <- cbind(segments_begin, segments_end)
       # Choose random segments for prediction
lib <- matrix(NA, nrow = length(segments_begin), ncol = 2)
lib[,1] <- segments_begin 
lib[,2] <- segments_end

pred <- lib
```

## Get embeddings from simplex

```{r}
simplex_out <- lapply(names(Imnaha)[3:11], function(var) {
simplex(Imnaha[, c("yr", var)], E = 2:10, lib = lib, pred = pred)
})

names(simplex_out) <- names(Imnaha)[-c(1,2)]

best_E <- sapply(simplex_out, function(df) {df$E[which.max(df$rho)]})
best_E

```

## Test non-linearity with S-map

```{r}

smap_out <- lapply(names(Imnaha)[3:11], function(var) {
  s_map(Imnaha[, c("yr", var)], E = best_E[var], lib = lib, 
        pred = pred)
})

names(smap_out) <- names(simplex_out)

par(mar = c(4, 4, 1, 1), mfrow = c(3, 3), mgp = c(2.5, 1, 0))  
# set up margins for plotting

lapply(names(smap_out), function(var) {
  plot(smap_out[[var]]$theta, smap_out[[var]]$rho, type = "l", xlab = "Nonlinearity (theta)", 
       ylab = "Forecast Skill (rho)", main = var)
})

```

## Univariate and Multivariate Forecast Models

```{r}
## Really need to write a function that lags variables automatically

data_by_stk <- split(Imnaha, Imnaha$stk)

lag_dat <- do.call(rbind, lapply(data_by_stk, function(df) {
  n <- NROW(df)
  temp <- data.frame(Year = df$yr)
  
  temp$rec3_t <- df$rec3_n
  temp$rec3_tm1 <- c(NA, temp$rec3_t[-n])
  temp$rec3_tm2 <- c(NA, temp$rec3_tm1[-n])
  
  temp$rec4_t <- df$rec4_n
  temp$rec4_tm1 <- c(NA, temp$rec4_t[-n])
  temp$rec4_tm2 <- c(NA, temp$rec4_tm1[-n])
  temp$rec4_tm3 <- c(NA, temp$rec4_tm2[-n])
  temp$rec4_tm4 <- c(NA, temp$rec4_tm3[-n])
  temp$rec4_tm5 <- c(NA, temp$rec4_tm4[-n])  
  temp$rec4_tm6 <- c(NA, temp$rec4_tm5[-n])
  
  temp$rec5_t <- df$rec5_n
  temp$rec5_tm1 <- c(NA, temp$rec5_t[-n])
  temp$rec5_tm2 <- c(NA, temp$rec5_tm1[-n])
 
  temp$eff_t <- df$eff_n
  temp$eff_tm1 <- c(NA, temp$eff_t[-n])
  temp$eff_tm2 <- c(NA, temp$eff_tm1[-n])

  temp$apr_t <- df$up48_apr
  temp$apr_tm1 <- c(NA, temp$apr_t[-n])
  temp$apr_tm2 <- c(NA, temp$apr_tm1[-n])
  
  temp$oct_t <- df$up48_oct
  temp$oct_tm1 <- c(NA, temp$oct_t[-n])
  temp$oct_tm2 <- c(NA, temp$oct_tm1[-n])
  
  temp$nov_t <- df$up48_nov
  temp$nov_tm1 <- c(NA, temp$nov_t[-n])
  temp$nov_tm2 <- c(NA, temp$nov_tm1[-n])
  
  temp$pdo_t <- df$pdo_win
  temp$pdo_tm1 <- c(NA, temp$pdo_t[-n])
  temp$pdo_tm2 <- c(NA, temp$pdo_tm1[-n])

  return(temp)
}))

head(lag_dat)
```


```{r}
univ_cols <- c("rec4_t", "rec4_tm1", "rec4_tm2", "rec4_tm3", "rec4_tm4", "rec4_tm5")
univ_output <- block_lnlp(lag_dat, lib = lib, pred = pred,columns = univ_cols, target_column = 1, stats_only = FALSE, first_column_time = TRUE)

univ_model_output <- univ_output[[1]]$model_output 
univ_observed <- univ_model_output$obs
univ_predicted <- univ_model_output$pred

univ_output[[1]]$stats 
```


```{r}
names(lag_dat)
multi_cols <- c("rec4_t", "rec4_tm1", "rec4_tm2", "apr_t", "eff_t", "oct_t")
multi_output <- block_lnlp(lag_dat, lib = lib, pred = pred, columns = multi_cols , target_column = 1, stats_only = FALSE, first_column_time = TRUE)

multi_model_output <- multi_output[[1]]$model_output 

multi_observed <- multi_model_output$obs
multi_predicted <- multi_model_output$pred

multi_output[[1]]$stats
```


Next, we run the models, being sure to set up predictor variables (the columns argument) and the variable to be predicted (the target_column argument). By default, predictions are always for one step ahead. Note that some models include only lagged observations of the target variable, and other models also include environmental data


```{r}
block_lnlp_4 <- function(block, target_column, columns, norm = FALSE)
{
    if(norm)
    {
        block[,columns] <- normalize(block[,columns])
    }
    
    lib_segments <- matrix(NA, nrow = 4, ncol = 2)
    segment_size <- NROW(block)/4
    start_index <- 1
    for(i in 1:4)
    {
        lib_segments[i,1] <- floor(start_index)
        end_index <- start_index - 1 + segment_size
        lib_segments[i,2] <- floor(end_index)
        start_index <- end_index+1
    }
    
    if(is.list(columns))
    {
        preds <- lapply(1:length(columns), function(x) {rep.int(NA, times = NROW(block))})
        for(i in 1:4)
        {
            pred_index <- lib_segments[i,1]:lib_segments[i,2]
            
            temp <- block_lnlp(block, lib = lib_segments[-i,], pred = lib_segments[i,], 
                               target_column = target_column, tp = 0, 
                               first_column_time = TRUE, 
                               columns = columns, stats_only = FALSE)
            
            for(j in 1:length(columns))
                preds[[j]][pred_index] <- temp$model_output[[j]]$pred[pred_index] #AI: changed from temp[[j]]$model_output$pred[pred_index]
        }
    }
    else
    {
        preds <- rep.int(NA, times = NROW(block))
        for(i in 1:4)
        {
            pred_index <- lib_segments[i,1]:lib_segments[i,2]
            
            temp <- block_lnlp(block, lib = lib_segments[-i,], pred = lib_segments[i,], 
                               target_column = target_column, tp = 0, 
                               first_column_time = TRUE, 
                               columns = columns, stats_only = FALSE)
            
            preds[pred_index] <- temp$model_output[[1]]$pred[pred_index] #AI: changed from temp[[1]]$model_output$pred[pred_index]

        }
    }
    return(preds)
}


simple_EDM <- function()
{
    forecast <- function(stock_df)
    {  
        make_forecasts <- function(block, mu_3, sigma_3, mu_4, sigma_4, mu_5, sigma_5)
        {
            rec3 <- block_lnlp_4(block, target_column = 2, columns = 1) #use function block_lnlp_4
            rec4 <- block_lnlp_4(block, target_column = 3, columns = 1)
            rec5 <- block_lnlp_4(block, target_column = 4, columns = 1)
            
            rec3 <- rec3*sigma_3 + mu_3
            rec4 <- rec4*sigma_4 + mu_4
            rec5 <- rec5*sigma_5 + mu_5
            return(rec3 + c(NA, rec4[1:(NROW(block)-1)]) + c(NA, NA, rec5[1:(NROW(block)-2)]))
        }
        
        # set up recruits and spawners
        valid <- is.finite(stock_df$rec) & is.finite(stock_df$eff) #exclude NAs? Should we??
        returns <- stock_df$ret[valid] #NAs left in here...
        years <- stock_df$yr[valid]
        spawners <- stock_df$eff_n[valid]
        recruits_3 <- stock_df$rec3_n[valid]
        mu_3 <- stock_df$rec3_mu[valid]
        sigma_3 <- stock_df$rec3_sigma[valid]
        recruits_4 <- stock_df$rec4_n[valid]
        mu_4 <- stock_df$rec4_mu[valid]
        sigma_4 <- stock_df$rec4_sigma[valid]
        recruits_5 <- stock_df$rec5_n[valid]
        mu_5 <- stock_df$rec5_mu[valid]
        sigma_5 <- stock_df$rec5_sigma[valid]
        
        # make block
        block <- data.frame(years = years, eff = spawners, 
                            rec3=recruits_3, rec4 = recruits_4, rec5 = recruits_5)
        
        if(length(returns) < 2) # check for enough data
            return(data.frame(year = NaN, obs = NaN, pred = NaN))
        
        forecasts <- make_forecasts(block,  mu_3, sigma_3, mu_4, sigma_4, mu_5, sigma_5)
        return(data.frame(year = years, obs = returns, pred = forecasts))
    }
    
    load("Data/Rdata/block_data.Rdata")
    
    # make forecasts for each stock
    results <- lapply(names(block_data), 
                      function(stk_name) {
                          cat("forecasting for ", stk_name, "... ", sep = "")
                          start_time <- proc.time()
                          output <- forecast(block_data[[stk_name]])
                          elapsed_time <- proc.time() - start_time
                          cat("(", elapsed_time[3], " sec.)\n", sep = "")
                          return(output)
                      })
    names(results) <- names(block_data)
    saveRDS(results, file = "Output/Rdata/results_simple_EDM.RDS")
    
    # compute stats
    stats <- do.call(rbind, lapply(results, function(stock_results) {
        compute_stats(stock_results$obs, stock_results$pred)
    }))
    stats$stk <- names(block_data)
    saveRDS(stats, file = "Output/Rdata/stats_simple_EDM.RDS")
    return()
}

simple_EDM()

```