---
title: "CCM_reduce_revised"
author: "Kurt Ingeman"
date: "6/5/2021"
output:
  html_document: default
  pdf_document: default
---


```{r}

rm(list = ls())

library(dplyr)
library(here)

```

## Read in 99 data and rename
```{r}

df <- read.csv(here("CORE EDM Visual", "CORE_reduce_CCM_vars", "CCM_twin_99.csv"))
# 258 vars

# ensure that max will work when all values negative
df[is.na(df)] <- -99

# split df into the rec4 and rec5 values
df4 <- df %>% 
  select(1:8) %>% 
  rename(
    ESU = ESU_rec4n,
    IMN = Imnaha_rec4n,
    MFS = Middle.Fork.Salmon_rec4n, 
    UPS = Upper.Salmon_rec4n
  )

df5 <- df %>% 
  select(1:4, 9:12) %>% 
  rename(
    ESU = ESU_rec5n,
    IMN = Imnaha_rec5n,
    MFS = Middle.Fork.Salmon_rec5n, 
    UPS = Upper.Salmon_rec5n
  )


```
## Rec4 -- Create df with only those variables x offset that represent the highest rho for at least one of the spatial components (ESU or one of the three MPGs)
```{r}

df4_cat <- df4 %>% 
  group_by(cat) %>% 
  mutate(cat_ESU = case_when(ESU > 0 & ESU == max(ESU) ~ 1)) %>% 
  mutate(cat_MFS = case_when(MFS > 0 & MFS == max(MFS) ~ 1)) %>%
  mutate(cat_IMN = case_when(IMN > 0 & IMN == max(IMN) ~ 1)) %>%
  mutate(cat_UPS = case_when(UPS> 0 & UPS == max(UPS) ~ 1)) %>% 
  ungroup() %>% rowwise() %>% 
  mutate(sum_cat = sum(cat_ESU, cat_MFS, cat_IMN, cat_UPS, na.rm = T))

#rec4 variables that represents the highest uRho in at least one full category: 
colSums(df4_cat[,9:12], na.rm = TRUE) 
#top_cat_ESU top_cat_MFS top_cat_IMN top_cat_UPS 
#         10           9           6           6 


rec4_top_cat <- df4_cat %>% 
  filter(sum_cat > 0)
# 28 variables

# write.csv(rec4_top_cat, "rec4_top_cat99.csv")
  

```

# Rec4 -- Do the same for subcategories
```{r}

df4_subcat <- df4 %>% 
  group_by(subcat) %>% 
  mutate(subcat_ESU = case_when(ESU > 0 & ESU == max(ESU) ~ 1)) %>% 
  mutate(subcat_MFS = case_when(MFS > 0 & MFS == max(MFS) ~ 1)) %>%
  mutate(subcat_IMN = case_when(IMN > 0 & IMN == max(IMN) ~ 1)) %>%
  mutate(subcat_UPS = case_when(UPS> 0 & UPS == max(UPS) ~ 1)) %>% 
  ungroup() %>% rowwise() %>% 
  mutate(sum_subcat = sum(subcat_ESU, subcat_MFS, subcat_IMN, subcat_UPS, na.rm = T))

#rec4 variables that represents the highest uRho in at least one  subcategory: 
colSums(df4_subcat[,9:12], na.rm = TRUE) 
#subcat_ESU    subcat_MFS   subcat_IMN  subcat_UPS 
#         46           35           11           8 


rec4_top_subcat <- df4_subcat %>% 
  filter(sum_subcat > 0)
# 70 variables

write.csv(rec4_top_subcat, "rec4_top_subcat99.csv")

```


## Rec 5 -- Create df with only those variables x offset that represent the highest rho for at least one of the spatial components (ESU or one of the three MPGs)
```{r}

df5_cat <- df5 %>% 
  group_by(cat) %>% 
  mutate(cat_ESU = case_when(ESU > 0 & ESU == max(ESU) ~ 1)) %>% 
  mutate(cat_MFS = case_when(MFS > 0 & MFS == max(MFS) ~ 1)) %>%
  mutate(cat_IMN = case_when(IMN > 0 & IMN == max(IMN) ~ 1)) %>%
  mutate(cat_UPS = case_when(UPS> 0 & UPS == max(UPS) ~ 1)) %>% 
  ungroup() %>% rowwise() %>% 
  mutate(sum_cat = sum(cat_ESU, cat_MFS, cat_IMN, cat_UPS, na.rm = T))

#rec4 variables that represents the highest uRho in at least one full category: 
colSums(df5_cat[,9:12], na.rm = TRUE) 
# cat_ESU cat_MFS cat_IMN cat_UPS 
#     11      10      11      10 

rec5_top_cat <- df5_cat %>% 
  filter(sum_cat > 0)
# 35 variables

write.csv(rec5_top_cat, "rec5_top_cat99.csv")
  

```

# Rec 5 -- Do the same for subcategories
```{r}

df5_subcat <- df5 %>% 
  group_by(subcat) %>% 
  mutate(subcat_ESU = case_when(ESU > 0 & ESU == max(ESU) ~ 1)) %>% 
  mutate(subcat_MFS = case_when(MFS > 0 & MFS == max(MFS) ~ 1)) %>%
  mutate(subcat_IMN = case_when(IMN > 0 & IMN == max(IMN) ~ 1)) %>%
  mutate(subcat_UPS = case_when(UPS> 0 & UPS == max(UPS) ~ 1)) %>% 
  ungroup() %>% rowwise() %>% 
  mutate(sum_subcat = sum(subcat_ESU, subcat_MFS, subcat_IMN, subcat_UPS, na.rm = T))

#rec5 variables that represents the highest uRho in at least one  subcategory: 
colSums(df5_subcat[,9:12], na.rm = TRUE) 
# subcat_ESU subcat_MFS subcat_IMN subcat_UPS 
#        54         48         37         38 

rec5_top_subcat <- df5_subcat %>% 
  filter(sum_subcat > 0)
# 107 variables

write.csv(rec5_top_subcat, "rec5_top_subcat99.csv")

```

## Figure out how many of those variable by offset combos are shared vs unique

```{r}
cat4 <- rec4_top_cat$FP
scat4 <- rec4_top_subcat$FP
cat5 <- rec5_top_cat$FP
scat5 <- rec5_top_subcat$FP

intersect(cat4,scat4) # 21 in common
intersect(cat4,cat5) # 6 in common
intersect(cat4,scat5) # 12 in common
intersect(scat4,cat5) # 13 in common
intersect(scat4,scat5) # 47 in common
intersect(cat5,scat5) # 30 in common 

Reduce(intersect, list(cat4, cat5, scat4, scat5))
# 2 in common to all
# "npgo.spr.2" and "orca.SRKWdeathsJKL.5"


vars99 <- list(cat4, scat4, cat5, scat5, names = c("cat4", "cat5", "scat4", "scat5"))

names(vars99) <- c("cat4", "scat4", "cat5", "scat5")

saveRDS(vars99, file = "vars99.rds")

```

## - 
## Repeat entire procedure for 95% CI
## -

```{r}

rm(list = ls())

library(dplyr)
library(here)

```

## Read in 95 data and rename
```{r}

df <- read.csv(here("CCM_twin_95.csv"))
# 250 vars

# ensure that max will work when all values negative
df[is.na(df)] <- -99

# split df into the rec4 and rec5 values
df4 <- df %>% 
  select(1:8) %>% 
  rename(
    ESU = ESU_rec4n,
    IMN = Imnaha_rec4n,
    MFS = Middle.Fork.Salmon_rec4n, 
    UPS = Upper.Salmon_rec4n
  )

df5 <- df %>% 
  select(1:4, 9:12) %>% 
  rename(
    ESU = ESU_rec5n,
    IMN = Imnaha_rec5n,
    MFS = Middle.Fork.Salmon_rec5n, 
    UPS = Upper.Salmon_rec5n
  )


```
## Rec4 -- Create df with only those variables x offset that represent the highest rho for at least one of the spatial components (ESU or one of the three MPGs)
```{r}

df4_cat <- df4 %>% 
  group_by(cat) %>% 
  mutate(cat_ESU = case_when(ESU > 0 & ESU == max(ESU) ~ 1)) %>% 
  mutate(cat_MFS = case_when(MFS > 0 & MFS == max(MFS) ~ 1)) %>%
  mutate(cat_IMN = case_when(IMN > 0 & IMN == max(IMN) ~ 1)) %>%
  mutate(cat_UPS = case_when(UPS> 0 & UPS == max(UPS) ~ 1)) %>% 
  ungroup() %>% rowwise() %>% 
  mutate(sum_cat = sum(cat_ESU, cat_MFS, cat_IMN, cat_UPS, na.rm = T))

#rec4 variables that represents the highest uRho in at least one full category: 
colSums(df4_cat[,9:12], na.rm = TRUE) 
#cat_ESU cat_MFS cat_IMN cat_UPS 
#     10      10      10      10 


rec4_top_cat <- df4_cat %>% 
  filter(sum_cat > 0)
# 35 variables

write.csv(rec4_top_cat, "rec4_top_cat95.csv")
  

```

## Create RDS for Alison with top cat by spatial segment
```{r}
rec4_vars_ESU <- df4_cat %>% 
  filter(cat_ESU > 0)
rec4_vars_ESU <- rec4_vars_ESU$FP

rec4_vars_MFS <- df4_cat %>% 
  filter(cat_MFS > 0)
rec4_vars_MFS <- rec4_vars_MFS$FP

rec4_vars_IMN <- df4_cat %>% 
  filter(cat_IMN > 0)
rec4_vars_IMN <- rec4_vars_IMN$FP

rec4_vars_UPS <- df4_cat %>% 
  filter(cat_UPS > 0)
rec4_vars_UPS <- rec4_vars_UPS$FP

vars_rec4_95 <- list(rec4_vars_ESU, rec4_vars_MFS, rec4_vars_IMN, rec4_vars_UPS)
names(vars_rec4_95) <- c("ESU", "MFS", "IMN", "UPS")


saveRDS(vars_rec4_95, file = "vars95_rec4_by_spatial.rds")

```

# Rec4 -- Do the same for subcategories
```{r}

df4_subcat <- df4 %>% 
  group_by(subcat) %>% 
  mutate(subcat_ESU = case_when(ESU > 0 & ESU == max(ESU) ~ 1)) %>% 
  mutate(subcat_MFS = case_when(MFS > 0 & MFS == max(MFS) ~ 1)) %>%
  mutate(subcat_IMN = case_when(IMN > 0 & IMN == max(IMN) ~ 1)) %>%
  mutate(subcat_UPS = case_when(UPS> 0 & UPS == max(UPS) ~ 1)) %>% 
  ungroup() %>% rowwise() %>% 
  mutate(sum_subcat = sum(subcat_ESU, subcat_MFS, subcat_IMN, subcat_UPS, na.rm = T))

#rec4 variables that represents the highest uRho in at least one  subcategory: 
colSums(df4_subcat[,9:12], na.rm = TRUE) 
#subcat_ESU subcat_MFS subcat_IMN subcat_UPS 
#        49         43         23         20 


rec4_top_subcat <- df4_subcat %>% 
  filter(sum_subcat > 0)
# 90 variables

write.csv(rec4_top_subcat, "rec4_top_subcat95.csv")

```

## Rec 5 -- Create df with only those variables x offset that represent the highest rho for at least one of the spatial components (ESU or one of the three MPGs)
```{r}

df5_cat <- df5 %>% 
  group_by(cat) %>% 
  mutate(cat_ESU = case_when(ESU > 0 & ESU == max(ESU) ~ 1)) %>% 
  mutate(cat_MFS = case_when(MFS > 0 & MFS == max(MFS) ~ 1)) %>%
  mutate(cat_IMN = case_when(IMN > 0 & IMN == max(IMN) ~ 1)) %>%
  mutate(cat_UPS = case_when(UPS> 0 & UPS == max(UPS) ~ 1)) %>% 
  ungroup() %>% rowwise() %>% 
  mutate(sum_cat = sum(cat_ESU, cat_MFS, cat_IMN, cat_UPS, na.rm = T))

#rec4 variables that represents the highest uRho in at least one full category: 
colSums(df5_cat[,9:12], na.rm = TRUE) 
# cat_ESU cat_MFS cat_IMN cat_UPS 
#     11      10      11      10 

rec5_top_cat <- df5_cat %>% 
  filter(sum_cat > 0)
# 33 variables

write.csv(rec5_top_cat, "rec5_top_cat95.csv")
  
```


## Create RDS for Alison with top cat by spatial segment
```{r}
rec5_vars_ESU <- df5_cat %>% 
  filter(cat_ESU > 0)
rec5_vars_ESU <- rec5_vars_ESU$FP

rec5_vars_MFS <- df5_cat %>% 
  filter(cat_MFS > 0)
rec5_vars_MFS <- rec5_vars_MFS$FP

rec5_vars_IMN <- df5_cat %>% 
  filter(cat_IMN > 0)
rec5_vars_IMN <- rec5_vars_IMN$FP

rec5_vars_UPS <- df5_cat %>% 
  filter(cat_UPS > 0)
rec5_vars_UPS <- rec5_vars_UPS$FP

vars_rec5_95 <- list(rec5_vars_ESU, rec5_vars_MFS, rec5_vars_IMN, rec5_vars_UPS)
names(vars_rec5_95) <- c("ESU", "MFS", "IMN", "UPS")


saveRDS(vars_rec5_95, file = "vars95_rec5_by_spatial.rds")

```


# Rec 5 -- Do the same for subcategories
```{r}

df5_subcat <- df5 %>% 
  group_by(subcat) %>% 
  mutate(subcat_ESU = case_when(ESU > 0 & ESU == max(ESU) ~ 1)) %>% 
  mutate(subcat_MFS = case_when(MFS > 0 & MFS == max(MFS) ~ 1)) %>%
  mutate(subcat_IMN = case_when(IMN > 0 & IMN == max(IMN) ~ 1)) %>%
  mutate(subcat_UPS = case_when(UPS> 0 & UPS == max(UPS) ~ 1)) %>% 
  ungroup() %>% rowwise() %>% 
  mutate(sum_subcat = sum(subcat_ESU, subcat_MFS, subcat_IMN, subcat_UPS, na.rm = T))

#rec5 variables that represents the highest uRho in at least one  subcategory: 
colSums(df5_subcat[,9:12], na.rm = TRUE) 
# subcat_ESU subcat_MFS subcat_IMN subcat_UPS 
#        54         49         47         45 

rec5_top_subcat <- df5_subcat %>% 
  filter(sum_subcat > 0)
# 107 variables

write.csv(rec5_top_subcat, "rec5_top_subcat95.csv")

```

## Figure out how many of those variable by offset combos are shared vs unique

```{r}
cat4 <- rec4_top_cat$FP
scat4 <- rec4_top_subcat$FP
cat5 <- rec5_top_cat$FP
scat5 <- rec5_top_subcat$FP

vars95 <- list(cat4, scat4, cat5, scat5)

names(vars95) <- c("cat4", "scat4", "cat5", "scat5")

saveRDS(vars95, file = "vars95.rds")

intersect(cat4,scat4) # 26 in common
intersect(cat4,cat5) # 9 in common
intersect(cat4,scat5) # 14 in common
intersect(scat4,cat5) # 14 in common
intersect(scat4,scat5) # 54 in common
intersect(cat5,scat5) # 27 in common 

Reduce(intersect, list(cat4, cat5, scat4, scat5))
# 5 in common to all
# "hatch.all.1" + "npgo.spr.2" + "orca.SRKWdeathsJKL.5" + "ssl.COL.2" + "upw.tdmi.4"

readRDS(file = "vars95.rds")

```
